### 2014-01-16 Studies/Chicago_2012/Report/Chicago_Fire.tex
Add case study on Chicago fire from 2012

\documentclass[11pt,oneside]{book}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                                                 %
% The mathematical style of these documents follows                                               %
%                                                                                                 %
% A. Thompson and B.N. Taylor. The NIST Guide for the Use of the International System of Units.   %
%    NIST Special Publication 881, 2008.                                                          %
%                                                                                                 %
% http://www.nist.gov/pml/pubs/sp811/index.cfm                                                    %
%                                                                                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{../../../Bibliography/commoncommands}

% Rename chapter headings
\renewcommand{\chaptername}{Section}
\renewcommand{\bibname}{References}

% Math shortcuts
\renewcommand{\sb}[1]{_\mathrm{#1}}
\renewcommand{\C}{\mbox{C}}
\renewcommand{\H}{\mbox{H}}
\renewcommand{\O}{\mbox{O}}
\newcommand{\N}{\mbox{N}}

\begin{document}

\bibliographystyle{unsrt}
\pagestyle{empty}

\begin{minipage}[t][9in][s]{6.25in}

\headerB{
Simulation of the Dynamics of an \\
Attic Fire in a 2 1/2 Story Wood \\
Frame Residential Structure \\
Chicago, IL \\
}

\headerC{
\flushright{
Craig G. Weinschenk \\
Daniel Madrzykowski \\
Kristopher J. Overholt \\
{\em Fire Research Division \\
Engineering Laboratory \\
Gaithersburg, Maryland, USA} \\ }
}

\flushright{\today \\
}

\vfill

\flushright{
\includegraphics[width=2.in]{../../../Bibliography/nistident_flright_vec} \\[.3in]
}

\titlesigs

\end{minipage}

\newpage

\frontmatter

\pagestyle{plain}
\pagenumbering{roman}

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{Contents}
\tableofcontents

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{List of Tables}
\listoftables

\chapter{List of Acronyms}

\begin{tabbing}
\hspace{1.5in} \= \\
BC \> Battalion Chief \\
FDS \> Fire Dynamics Simulator \\
HGL \> Hot Gas Layer \\
HRR \> Heat Release Rate \\
NIST \> National Institute of Standards and Technology \\
SFPE \> Society of Fire Protection Engineers \\
\end{tabbing}

\mainmatter

\chapter{Introduction}

\chapter{Fire Incident Summary}
The following account of events is based on information provided by NIST physical measurements on-scene, NIST photographs of the scene, time line information provided by the Chicago Fire Department and information provided from the NIOSH Fire Fighter Fatality Investigation and Prevention Program Report \cite{NIOSH:Bowyer}.

On November 2, 2012, a fire in a 2 1/2 story wood frame residential structure resulted in the death of a death of a fire captain and injury to a fire fighter.  At approximately 5:16 in the afternoon, a dispatch was made for ``smoke in the area''.   Within a minute of dispatch, a Battalion Chief (BC) arrived and sized up the structure fire from the exterior and reported a working fire with heavy smoke coming from the front and rear of the structure's attic.  After speaking with the building occupants, the BC entered the structure and found light smoke on the second floor and a glow around the doorway that led from the second floor hallway to enclosed rear porch.  Engine 123 arrived on scene at approximately 5:19. The victims were part of the Engine 123 crew.  The BC met the Captain and the Fire Fighter (victims) at the front door to the structure.  Then the Captain and Fire Fighter took their hose line to the second floor of the structure.  At approximately 5:21, the Captain and Fire Fighter are on the second floor with their hoseline and had begun advancing the charged line toward the rear second floor door between the apartment and the enclosed porch.   Approximately 30 seconds later, a fire fighter from Tower Ladder 39 (TL 39) enters the enclosed porch on the rear of the structure by opening the rear exterior door to the first floor of the porch.   The fire fighter from TL 39, notices ``the fire light up'' in the stairwell of the enclosed porch, and then kicks in the first level door between the rear porch and the main structure.   Both doors remain open and the fire fighter from TL 39 exits the structure.  Given the amount of fire showing from the second floor of the rear porch and ``heavy fire'' in the stairwell of the rear porch, the BC radios the Engine 123 Captain at 5:23, to alert him to the fire conditions and to advise him that Engine 49 would be applying water to the fire on the rear of the structure from the exterior.  No response was made by the Engine 123 Captain.  It is believed that the door between the rear porch and the second floor hallway, where the Engine 123 Captain and Fire Fighter were located may have failed by this time.  Approximately a minute later Engine 49 applied water to the fire in the attic.

Squad 5 entered the front of the structure at approximately 5:25 and headed to the second floor.  They heard a commotion and found the Engine 123 Captain collapsed on the kitchen floor and entangled with the Engine 123 firefighter.  A Mayday was called on the radio by Squad 5 at approximately  5:27 and the Engine 123 Captain was removed from the front of the building at approximately 5:29 and succumbed to his injuries at the hospital at 5:38.

\begin{table}
\centering
\caption{Fire and Simulation Event Timeline}
\vskip.5\baselineskip
\begin{tabular}{|c|c|c|}
\hline
Fire Time & Simulation Time & Action  \\
\hline \hline
        &        & Reported working fire with heavy smoke\\
5:17    &        & coming from the front and rear of the \\
        &        & structure's attic. \\ \hline
        &  0:00  & Simulation begins. Attic fire ignites. Initial \\ 
        &        & self ventilation on charlie side. \\ \hline
        &  0:20  & Front door opens. \\ \hline
5:20    &  0:40  & Attic self ventilation - heavy smoke reported. Porch fire ignites. \\ \hline
        &  0:45  & Additional self ventilation on charlie side. \\ \hline
        &  0:55  & Self ventilation on beta side. \\ \hline        
5:21:00 &        & Engine 123 Capt. and FF on 2$^{nd}$ floor. \\ \hline
        &        & Rear Exterior $-$ 1$^{st}$ floor door to enclosed \\
5:21:30 &  2:10  & porch opened, fire in rear porch begins to \\
        &        & increase HRR. \\ \hline
        &  2:15  & Additional self ventilation on beta side. \\ \hline        
        &  2:40  & Top of interior door fails. \\ \hline
5:23    & $\approx$ 3:40       & No response from Engine 123: Capt. or FF. \\ \hline
5:24    &  N/A   & Water applied to attic from the rear exterior position. \\ \hline
        &  5:00  & Simulation ends. \\ \hline
\end{tabular}
\label{tab:fires}
\end{table}

\chapter{Fire Dynamics Simulator and Smokeview}
FDS is a computational fluid dynamics (CFD) model that solves a form of the Navier-Stokes equations appropriate for low-speed, thermally driven flow with an emphasis on smoke and heat transport from fires.  Within a CFD model, the room or building is divided into small three-dimensional rectangular control volumes or computational cells.  The cells are contained together within one larger volume known as a computational domain.  The CFD model computes the density, velocity, temperature, pressure and species concentration of the gas in each cell.  Based on the laws of conservation of mass, momentum,  and energy, the model tracks the generation and movement of fire gasses.  FDS can utilize material properties of the furnishings, walls, floors, and ceilings to compute fire growth and spread.  One of the most important advantages of FDS is that it is verified and validated against fire test data to ensure that it provides the expected results, given sufficient input data \cite{FDS_Verification_Guide,FDS_Validation_Guide}.  A complete description of the FDS model is given in reference \cite{FDS_Tech_Guide}.

Smokeview is a software tool designed to visualize the results of the FDS \cite{Smokeview_Users_Guide}.  Smokeview visualizes smoke and other attributes of the modeled fire using traditional scientific methods such as displaying tracer particle flow, two dimensional (92D) or three dimensional (3D) shaded contours of gas flow data such as temperature and flow vectors showing flow direction and magnitude.  Smokeview allows the fire and fire movement to be visualized.  This is done by displaying a series of partially transparent planes where the transparencies in each plane (at each grid node) are determined from soot densities computed by FDS.  Smokeview also visualizes static data at particular times using 2D or 3D contours of data such as temperature and flow vectors showing flow direction and magnitude.

\chapter{Model Input Parameters}
Input data from a number of sources must be collected and documented in order to simulate a fire with FDS or any other model. 
From the fire scene the following information must be gathered:
\begin{itemize}
\item The geometry of the building and the compartments being modeled.
\item The size and location of exterior and interior ventilation openings. 
\item Documentation and identification of the fuels used in the construction and finish of the building and their installed location
\item Documentation and identification of the fuels used as the contents of the building. 
\item Area of origin and approximate size of ignition source.
\item Documentation of fire damage to the building, fuel loads, victims or equipment.
\end{itemize}
From witnesses, first responders, reports, and recorded media such as fire ground radios or videos:
\begin{itemize}
\item Information on the timing of the fire development
\item Order and timing of ventilation opening to the outside
\item Weather conditions at the time of the fire
\end{itemize}
To develop the simulation, a timeline of events based on the collected information must be developed.   From this time line actions may be input into the model, such as venting a window at a given time.  Or the time line may be used in validating a model result.

The development of the model geometry and development of the input file is straight forward with guidance from the FDS users' manual \cite{FDS_Users_Guide} or third party software.  Engineering judgment, graduate level knowledge of fire dynamics, and modeling experience are required to  translate the information collected from the scene into appropriate and realistic computational information.  Examples of these requirements include:
\begin{itemize}
\item the computational cell size
\item model domain and boundaries conditions
\item the heat release rate (HRR) of the ignition source/initial fire
\item physical and thermal properties of fuels such as interior finish and furnishings
	\begin{itemize}
	\item density
	\item specific heat
	\item thermal conductivity
	\item ignition temperature
	\item heat of combustion
	\item heat release rate
	\item mass loss rate under various fire conditions
	\item pyrolyzate/combustion product emissions
	\item response to thermal exposures
	\item activation energy
	\end{itemize}
\end{itemize}
Since there is no all-inclusive, universal fire modeling input data base this type of information must be gathered from literature or generated for a given incident.  The sections below describe the input data and how it was modeled.

\section{Geometry}
\label{geom}
The structure is a 2 $\tfrac{1}{2}$ story residential structure with balloon frame construction and a wood gable roofing system. The structure has a 17.15 m (56 ft) by 6.15 m (20 ft) footprint. This results in a floor area of approximately 105 m$^2$ (1100 ft$^2$) per floor and a total living area of approximately 110 m$^2$ (2200 ft$^2$). Additionally there is a half-story attic that was used to house the structure's hot water heater. Interior walls were constructed from wood frame dimensioned lumber with 1.27 cm (0.5 in) gypsum board over top. There were two stairwells in the structure: the front stairwell connected the first and second levels and the rear stairwell (part of the enclosed porch) connected the first and second levels and provided access to the attic. Figs.~(\ref{fig:first_floor} and \ref{fig:second_floor}) show the interior dimensions of the first and second floor, respectively. The dimensions were found from on-scene measurements using a tape measure. Measurements were made to the nearest 1.27 cm (0.5 in). Based on difficulty of making some of the measurements due to the condition of the post-fire structure, the uncertainty in the measurements is estimated to be on the order of $\pm$ 1 \%. This would result in measurements that could change by up $\pm$ 17 cm ($\approx$ 7 in). When considering the uncertainties in the fire ignition time, ventilation, and peak heat release rate among other factors, the uncertainty in the overall structure geometry is not expected to be a major factor.

The model geometry was constructed based on the measurements taken at the structure. The resolution of the model is to the nearest centimeter with that the recognition that the resolution is ultimately limited by the mesh resolution which will be discussed in Section \ref{mesh}. Fig.~(\ref{fig:smv_interior}) shows Smokeview renderings of cutaways of the beta and delta sides of the structure interior.

\begin{figure}[h!]
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
\includegraphics[width=.80\textwidth]{../Figures/interior_beta} \\
\includegraphics[width=.80\textwidth]{../Figures/interior_delta} \\
\end{tabular*}
\caption{3D Smokeview rendering of structure interior looking from the charlie side: to the beta side (top) and to the delta side (bottom)}
\label{fig:smv_interior}
\end{figure}

\section{Fire}
\label{fire}

To estimate the dominant fuel(s), fire size, and structure ventilation, on-scene videos and post-incident reports are used. The fire was first noticed by a person on the street who informed the residents that smoke was pouring out of the attic. The residents, upon notification of the fire, observed that there was fire in their enclosed rear porch. Post-incident, the fire investigation concluded that the fire originated in the structure's attic.

Based on these findings, the FDS model includes two source fire locations: an initial fire in the attic and a secondary fire on the second floor porch landing. The attic and rear porch consist of a fuel load that is mainly wood: the rear porch was enclosed with wood planks covered by foam board and vinyl siding, the roof was asphalt shingles over wood planking, and the attic had oriented strand board (OSB) floors. If we represent wood with the following chemical formula, $\C\H_{1.7}\O_{0.74}\N_{0.002}$, and use carbon monoxide and soot yields from the SFPE Handbook \cite{SFPE:Tewarson}, we can write a balanced chemical reaction for wood oxidation:

\begin{multline}
\C\H_{1.7}\O_{0.74}\N_{0.002} + 4.86(0.21\,\O_{2} + 0.79\,\N_{2}) \\ 
\rightarrow 5.69(0.0006\,\C\O + 0.17\,\C\O_{2} + 0.15\,\H_{2}\O + 0.674\,\N_{2} + 0.006\,\C)
\label{eq:wood_comb}
\end{multline}
The heat of combustion of wood for this simulation is 16400 kJ/kg (7050 BTU/lb), based on data provided in the SFPE Handbook \cite{SFPE:Tewarson}. The heat of combustion quantifies the energy release per unit mass of the fuel.

From work by Babrauskas and Grayson \cite{babrauskas1990}, we can find an estimate for the energy release per unit time and unit area (kW/m$^2$). Experiments were conducted in a cone calorimeter to determine the 5 minute average of HRR (kW/m$^2$) for several different types of wood over a range of radiant heat fluxes (kW/m$^2$). Taking a conservative approach, we assume the burning wood in the attic had the lowest energy density of the tested samples \cite{babrauskas1990}, similar to that of Douglas Fir plywood. If we estimate that the average radiation feedback to the pyrolyzing fuel is 30 kW/m$^2$ (assuming under-ventilated conditions within the structure) then an average heat release for wood is 65 kW/m$^2$.

Next, is determining the appropriate burnable area (exposed wood area) in the attic and in the second floor porch enclosure. Based on the floor plan of the structure, the underside of the roof is approximately 134 m$^2$ (1475 ft$^2$) plus there is 105 m$^2$ of floor area in the attic. Examination of post incident images showed the roof sustained more damage than the floor except for the floor area immediately over the back porch. That area will be considered as part of the porch fire. To estimate the attic fire size, we consider the area of the underside of the roof (134 m$^2$) and use an average heat release rate of wood of 65 kW/m$^2$. This results in a fire size of about 9 MW. To estimate the porch fire size, we consider the the four walls of porch enclosure (42 m$^2$) and the ceiling above the porch (13 m$^2$). The total area is 55 m$^2$ (592 ft$^2$). Using the same average heat release of 65 kW/m$^2$, the fire size on the porch is approximately 3.3 MW. The total fire size for this structure is estimated to be 12.3 MW.

Since it is unknown when the fire started relative to notification of a fire, the attic fire starts at the beginning of the simulation and ramps up to 9 MW in under 10 seconds. At the same time, part of the second floor window on the charlie side, 0.11 m$^2$ (1.2 ft$^2$) , fails providing initial ventilation.  The attic fire starts 40 seconds into the simulation and ramps to 3.3 MW in under 10 seconds. This corresponds to the report of heavy smoke from both the alpha, 0.08 m$^2$ (0.86 ft$^2$), and charlie, 0.1 m$^2$ (1 ft$^2$), sides of the attic. Additional ventilation occurs 45 and 55 seconds into the simulation as the fire burns through the exterior wall, 0.12 m$^2$ (1.3 ft$^2$), on both the beta and charlie side of the structure. At 130 seconds into the simulation, the back door is opened by a firefighter which results in a final fire induced ventilation. A hole, 0.2 m$^2$ (2.2 ft$^2$),  opens on the rear porch of the beta side at 135 seconds.

\section{Mesh}
\label{mesh}

For the simulation in this study, we can estimate a measure of how well the flow field is resolved by the non-dimensional expression $D^*/\dx$. Here $D^*$ is a characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g} }
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}   
From the FDS User Guide \cite{FDS_Users_Guide}, ``the characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$.'' $D$ is the physical diameter of the fire. Following Eq.~(\ref{eq:mesh}) and using 10 cm grid cells, the characteristic fire diameter to cell size ratio based on the 3.3 MW porch fire is 15.5 and for the 9 MW fire is 23. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16 \cite{NUREG_1824}. The grid resolution used in the model fall within typical, engineering $D^*/\dx$ values.

\chapter{Model Results}

\section{Heat Release Rate}
\label{HRR}
The heat release rate output data from FDS quantifies the model-predicted rate of energy release in the simulation. Examination of this information can indicate ignition time, if a fire is under ventilated (FDS HRR less than prescribed HRR), the impact of ventilation, or if complete extinction occurs among other characteristics of fire behavior. Fig.~(\ref{fig:hrr}) shows the HRR versus time based on the defined input fires described in Section \ref{fire}. The two vertical lines represent when a firefighter opened the back door of the structure (dotted) and when the second floor interior door was estimated to fail (dash-dot).

\begin{figure}[h!]
\centering
\includegraphics[width=.80\textwidth]{../Figures/Chicago_Fire_HRR}
\caption{HRR versus time of FDS model of the West 50th St. fire. The first vertical line indicates when the back door failed, the second indicates when the interior hallway door failed.}
\label{fig:hrr}
\end{figure}

Just prior to the back door being opened, the HRR was 6.6 MW. This is below the specified HRR of 12.3 MW, indicated that the fire was under-ventilated. From the time the back doored was opened to estimated time the interior door failed (approximately 30 seconds), the HRR increased by approximately 35 \% to 8.9 MW.

\section{Temperature}

\section{Velocity}

\section{Pressure}

\section{Gas Concentrations}

\chapter{Discussion}

\section{Conditions in Hallway}

\section{Comparison of conditions with PPE and Victim Limitations}

\section{Tactical Considerations}

\chapter{Summary}

\chapter{Acknowledgements}

\bibliography{../../../Bibliography/FDS_refs,../../../Bibliography/FDS_general,chicago}

\appendix

\chapter{Dimensioned Drawings}

\begin{figure}[h!]
\centering
\includegraphics[width=.80\textwidth]{../Figures/50th_Place_1st_Floor}
\caption {Dimensioned drawing of 1$^{st}$ floor}
\label{fig:first_floor}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.80\textwidth]{../Figures/50th_Place_2nd_Floor}
\caption {Dimensioned drawing of 2$^{nd}$ floor}
\label{fig:second_floor}
\end{figure}

\end{document}