### 2014-04-21 Projects/DelCo_2014_04/Report/DelCo_2014.tex
DelCo: Edits to report and timing labels

\documentclass[11pt,oneside]{book}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                                                 %
% The mathematical style of these documents follows                                               %
%                                                                                                 %
% A. Thompson and B.N. Taylor. The NIST Guide for the Use of the International System of Units.   %
%    NIST Special Publication 881, 2008.                                                          %
%                                                                                                 %
% http://www.nist.gov/pml/pubs/sp811/index.cfm                                                    %
%                                                                                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{../../../Bibliography/commoncommands}

% Rename chapter headings
\renewcommand{\chaptername}{Section}
\renewcommand{\bibname}{References}

% Math shortcuts
\renewcommand{\sb}[1]{_\mathrm{#1}}
\renewcommand{\C}{\mbox{C}}
\renewcommand{\H}{\mbox{H}}
\renewcommand{\O}{\mbox{O}}
\newcommand{\N}{\mbox{N}}

% Center all figures
\makeatletter
\g@addto@macro\@floatboxreset\centering
\makeatother

\begin{document}

\bibliographystyle{unsrt}
\pagestyle{empty}

\begin{minipage}[t][9in][s]{6.25in}

\headerB{
Delco 2014
}

\headerC{
\flushright{
Daniel Madrzykowski \\
Craig G. Weinschenk \\
Kristopher J. Overholt \\
{\em Fire Research Division \\
Engineering Laboratory \\
Gaithersburg, Maryland, USA} \\ }
}

\flushright{\today \\
}

\vfill

\flushright{
\includegraphics[width=2.in]{../../../Bibliography/nistident_flright_vec} \\[.3in]
}

\titlesigs

\end{minipage}

\newpage

\frontmatter

\pagestyle{plain}
\pagenumbering{roman}

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{Contents}
\tableofcontents

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{List of Tables}
\listoftables

\chapter{List of Acronyms}

\begin{tabbing}
\hspace{1.5in} \= \\
FDS \> Fire Dynamics Simulator \\
HGL \> Hot Gas Layer \\
HRR \> Heat Release Rate \\
HRRPUA \> Heat Release Rate per Unit Area \\
NIST \> National Institute of Standards and Technology \\
\end{tabbing}

\mainmatter

\chapter{Introduction}
\label{chap:Introduction}

\chapter{Experimental Setup}
\label{chap:Experimental_Setup}

\chapter{Results}
\label{chap:Results}

\section{Hose Flow Tests}

\subsection{DNTADM}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DNTADM_BDP_A6}
\caption{BDP at interior door CD, Test DNTADM}
\label{fig:HOSE_DNTADM_BDP_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DNTADM_BDP_A7}
\caption{BDP at exterior door A, Test DNTADM}
\label{fig:HOSE_DNTADM_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DNTADM_BDP_A8}
\caption{BDP at exterior door C\_West, Test DNTADM}
\label{fig:HOSE_DNTADM_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DNTADM_BDP_A9}
\caption{BDP at exterior door C\_East, Test DNTADM}
\label{fig:HOSE_DNTADM_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DNTADM_HOSE}
\caption{Hose pressure, Test DNTADM}
\label{fig:HOSE_DNTADM_HOSE}
\end{figure}

\clearpage

\subsection{DNTODF}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DNTODF_BDP_A6}
\caption{BDP at interior door CD, Test DNTODF}
\label{fig:HOSE_DNTODF_BDP_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DNTODF_BDP_A7}
\caption{BDP at exterior door A, Test DNTODF}
\label{fig:HOSE_DNTODF_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DNTODF_BDP_A8}
\caption{BDP at exterior door C\_West, Test DNTODF}
\label{fig:HOSE_DNTODF_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DNTODF_BDP_A9}
\caption{BDP at exterior door C\_East, Test DNTODF}
\label{fig:HOSE_DNTODF_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DNTODF_HOSE}
\caption{Hose pressure, Test DNTODF}
\label{fig:HOSE_DNTODF_HOSE}
\end{figure}

\clearpage

\subsection{DSTADF}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DSTADF_BDP_A6}
\caption{BDP at interior door CD, Test DSTADF}
\label{fig:HOSE_DSTADF_BDP_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DSTADF_BDP_A7}
\caption{BDP at exterior door A, Test DSTADF}
\label{fig:HOSE_DSTADF_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DSTADF_BDP_A8}
\caption{BDP at exterior door C\_West, Test DSTADF}
\label{fig:HOSE_DSTADF_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DSTADF_BDP_A9}
\caption{BDP at exterior door C\_East, Test DSTADF}
\label{fig:HOSE_DSTADF_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DSTADF_HOSE}
\caption{Hose pressure, Test DSTADF}
\label{fig:HOSE_DSTADF_HOSE}
\end{figure}

\clearpage

\subsection{DSTODF}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DSTODF_BDP_A6}
\caption{BDP at interior door CD, Test DSTODF}
\label{fig:HOSE_DSTODF_BDP_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DSTODF_BDP_A7}
\caption{BDP at exterior door A, Test DSTODF}
\label{fig:HOSE_DSTODF_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DSTODF_BDP_A8}
\caption{BDP at exterior door C\_West, Test DSTODF}
\label{fig:HOSE_DSTODF_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DSTODF_BDP_A9}
\caption{BDP at exterior door C\_East, Test DSTODF}
\label{fig:HOSE_DSTODF_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_DSTODF_HOSE}
\caption{Hose pressure, Test DSTODF}
\label{fig:HOSE_DSTODF_HOSE}
\end{figure}

\clearpage

\subsection{IMAODM}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_IMAODM_BDP_A6}
\caption{BDP at interior door CD, Test IMAODM}
\label{fig:HOSE_IMAODM_BDP_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_IMAODM_BDP_A7}
\caption{BDP at exterior door A, Test IMAODM}
\label{fig:HOSE_IMAODM_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_IMAODM_BDP_A8}
\caption{BDP at exterior door C\_West, Test IMAODM}
\label{fig:HOSE_IMAODM_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_IMAODM_BDP_A9}
\caption{BDP at exterior door C\_East, Test IMAODM}
\label{fig:HOSE_IMAODM_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_IMAODM_HOSE}
\caption{Hose pressure, Test IMAODM}
\label{fig:HOSE_IMAODM_HOSE}
\end{figure}

\clearpage

\subsection{IMTADM}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_IMTADM_BDP_A6}
\caption{BDP at interior door CD, Test IMTADM}
\label{fig:HOSE_IMTADM_BDP_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_IMTADM_BDP_A7}
\caption{BDP at exterior door A, Test IMTADM}
\label{fig:HOSE_IMTADM_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_IMTADM_BDP_A8}
\caption{BDP at exterior door C\_West, Test IMTADM}
\label{fig:HOSE_IMTADM_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_IMTADM_BDP_A9}
\caption{BDP at exterior door C\_East, Test IMTADM}
\label{fig:HOSE_IMTADM_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/HOSE/HOSE_IMTADM_HOSE}
\caption{Hose pressure, Test IMTADM}
\label{fig:HOSE_IMTADM_HOSE}
\end{figure}

\chapter{Conclusions}
\label{chap:Conclusions}

\chapter{Acknowledgements}
\label{chap:Acknowledgements}

\bibliography{../../../Bibliography/FDS_refs,../../../Bibliography/FDS_general,references}

\appendix

\chapter{Appendix A}

Placeholder

\end{document}