### 2014-07-04 Projects/DelCo_2014_04/Report/DelCo_2014.tex
DelCo: Move plots dir

\documentclass[11pt,oneside]{book}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                                                 %
% The mathematical style of these documents follows                                               %
%                                                                                                 %
% A. Thompson and B.N. Taylor. The NIST Guide for the Use of the International System of Units.   %
%    NIST Special Publication 881, 2008.                                                          %
%                                                                                                 %
% http://www.nist.gov/pml/pubs/sp811/index.cfm                                                    %
%                                                                                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{../../../Bibliography/commoncommands}

% Rename chapter headings
\renewcommand{\chaptername}{Section}
\renewcommand{\bibname}{References}

% Math shortcuts
\renewcommand{\sb}[1]{_\mathrm{#1}}
\renewcommand{\C}{\mbox{C}}
\renewcommand{\H}{\mbox{H}}
\renewcommand{\O}{\mbox{O}}
\newcommand{\N}{\mbox{N}}

% Center all figures
\makeatletter
\g@addto@macro\@floatboxreset\centering
\makeatother

\begin{document}

\bibliographystyle{unsrt}
\pagestyle{empty}

\begin{minipage}[t][9in][s]{6.25in}

\headerB{
Delco 2014
}

\headerC{
\flushright{
Daniel Madrzykowski \\
Craig G. Weinschenk \\
Kristopher J. Overholt \\
{\em Fire Research Division \\
Engineering Laboratory \\
Gaithersburg, Maryland, USA} \\ }
}

\flushright{\today \\
}

\vfill

\flushright{
\includegraphics[width=2.in]{../../../Bibliography/nistident_flright_vec} \\[.3in]
}

\titlesigs

\end{minipage}

\newpage

\frontmatter

\pagestyle{plain}
\pagenumbering{roman}

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{Contents}
\tableofcontents

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{List of Tables}
\listoftables

\chapter{List of Acronyms}

\begin{tabbing}
\hspace{1.5in} \= \\
FDS \> Fire Dynamics Simulator \\
HGL \> Hot Gas Layer \\
HRR \> Heat Release Rate \\
HRRPUA \> Heat Release Rate per Unit Area \\
NIST \> National Institute of Standards and Technology \\
\end{tabbing}

\mainmatter

\chapter{Introduction}
\label{chap:Introduction}

\chapter{Experimental Setup}
\label{chap:Experimental_Setup}

\section{Experimental Structure}
\label{sec:Experimental_Structure}

\subsection{Construction}
\label{sec:Construction}
The building was composed of concrete walls, as shown in Fig.~\ref{fig:EL_Logo_Example}.

\begin{figure}[!ht]
\includegraphics[width=2in]{../Figures/el_logo_small}
\caption{NIST Engineering Laboratory}
\label{fig:EL_Logo_Example}
\end{figure}

\section{Fuel Load}

\chapter{Results}
\label{chap:Results}

\section{Hose Flow Tests}

\subsection{DNXAXX}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DNXAXX_BDP_A7}
\caption{Velocity at exterior Door A, Test HOSE\_DNXAXX}
\label{fig:HOSE_DNXAXX_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DNXAXX_BDP_A8}
\caption{Velocity at exterior Door C\_West, Test HOSE\_DNXAXX}
\label{fig:HOSE_DNXAXX_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DNXAXX_BDP_A9}
\caption{Velocity at exterior Door C\_East, Test HOSE\_DNXAXX}
\label{fig:HOSE_DNXAXX_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DNXAXX_HOSE}
\caption{Hose pressure, Test HOSE\_DNXAXX}
\label{fig:HOSE_DNXAXX_HOSE}
\end{figure}

\clearpage

\subsection{DNXOXF}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DNXOXF_BDP_A7}
\caption{Velocity at exterior Door A, Test HOSE\_DNXOXF}
\label{fig:HOSE_DNXOXF_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DNXOXF_BDP_A8}
\caption{Velocity at exterior Door C\_West, Test HOSE\_DNXOXF}
\label{fig:HOSE_DNXOXF_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DNXOXF_BDP_A9}
\caption{Velocity at exterior Door C\_East, Test HOSE\_DNXOXF}
\label{fig:HOSE_DNXOXF_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DNXOXF_HOSE}
\caption{Hose pressure, Test HOSE\_DNXOXF}
\label{fig:HOSE_DNXOXF_HOSE}
\end{figure}

\clearpage

\subsection{DSXAXF}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DSXAXF_BDP_A7}
\caption{Velocity at exterior Door A, Test HOSE\_DSXAXF}
\label{fig:HOSE_DSXAXF_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DSXAXF_BDP_A8}
\caption{Velocity at exterior Door C\_West, Test HOSE\_DSXAXF}
\label{fig:HOSE_DSXAXF_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DSXAXF_BDP_A9}
\caption{Velocity at exterior Door C\_East, Test HOSE\_DSXAXF}
\label{fig:HOSE_DSXAXF_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DSXAXF_HOSE}
\caption{Hose pressure, Test HOSE\_DSXAXF}
\label{fig:HOSE_DSXAXF_HOSE}
\end{figure}

\clearpage

\subsection{DSXOXF}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DSXOXF_BDP_A7}
\caption{Velocity at exterior Door A, Test HOSE\_DSXOXF}
\label{fig:HOSE_DSXOXF_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DSXOXF_BDP_A8}
\caption{Velocity at exterior Door C\_West, Test HOSE\_DSXOXF}
\label{fig:HOSE_DSXOXF_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DSXOXF_BDP_A9}
\caption{Velocity at exterior Door C\_East, Test HOSE\_DSXOXF}
\label{fig:HOSE_DSXOXF_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_DSXOXF_HOSE}
\caption{Hose pressure, Test HOSE\_DSXOXF}
\label{fig:HOSE_DSXOXF_HOSE}
\end{figure}

\clearpage

\subsection{IXAOXX}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_IXAOXX_BDP_A6}
\caption{Velocity at interior Door CD, Test HOSE\_IXAOXX}
\label{fig:HOSE_IXAOXX_BDP_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_IXAOXX_BDP_A7}
\caption{Velocity at exterior Door A, Test HOSE\_IXAOXX}
\label{fig:HOSE_IXAOXX_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_IXAOXX_BDP_A8}
\caption{Velocity at exterior Door C\_West, Test HOSE\_IXAOXX}
\label{fig:HOSE_IXAOXX_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_IXAOXX_BDP_A9}
\caption{Velocity at exterior Door C\_East, Test HOSE\_IXAOXX}
\label{fig:HOSE_IXAOXX_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_IXAOXX_HOSE}
\caption{Hose pressure, Test HOSE\_IXAOXX}
\label{fig:HOSE_IXAOXX_HOSE}
\end{figure}

\clearpage

\subsection{IXXAXX}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_IXXAXX_BDP_A6}
\caption{Velocity at interior Door CD, Test HOSE\_IXXAXX}
\label{fig:HOSE_IXXAXX_BDP_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_IXXAXX_BDP_A7}
\caption{Velocity at exterior Door A, Test HOSE\_IXXAXX}
\label{fig:HOSE_IXXAXX_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_IXXAXX_BDP_A8}
\caption{Velocity at exterior Door C\_West, Test HOSE\_IXXAXX}
\label{fig:HOSE_IXXAXX_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_IXXAXX_BDP_A9}
\caption{Velocity at exterior Door C\_East, Test HOSE\_IXXAXX}
\label{fig:HOSE_IXXAXX_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/HOSE_IXXAXX_HOSE}
\caption{Hose pressure, Test HOSE\_IXXAXX}
\label{fig:HOSE_IXXAXX_HOSE}
\end{figure}

\clearpage

\section{Fire Tests}

\subsection{FIRE\_FA2S1C1NCX1\_1}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_BDP_A7}
\caption{Velocity at exterior Door A, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_TC_A7}
\caption{Temperature at exterior Door A, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_BDP_A8}
\caption{Velocity at exterior Door C\_West, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_TC_A8}
\caption{Temperature at exterior Door C\_West, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_TC_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_BDP_A9}
\caption{Velocity at exterior Door C\_East, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_TC_A9}
\caption{Temperature at exterior Door C\_East, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_TC_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_GAS}
\caption{Gas Concentration, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_GAS}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_HF}
\caption{Heat Flux, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_TC_A1}
\caption{Temperature at TC Tree A1, Room A, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_TC_A2}
\caption{Temperature at TC Tree A2, Room A, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_TC_A3}
\caption{Temperature at TC Tree A3, Room B, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_TC_A4}
\caption{Temperature at TC Tree A4, Room C, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_1_TC_A5}
\caption{Temperature at TC Tree A5, Room C, Test FIRE\_FA2S1C1NCX1\_1}
\label{fig:FIRE_FA2S1C1NCX1_1_TC_A5}
\end{figure}

\clearpage

\subsection{FIRE\_FA2S1C1NCX1\_2}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_BDP_A7}
\caption{Velocity at exterior Door A, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_TC_A7}
\caption{Temperature at exterior Door A, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_BDP_A8}
\caption{Velocity at exterior Door C\_West, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_TC_A8}
\caption{Temperature at exterior Door C\_West, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_TC_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_BDP_A9}
\caption{Velocity at exterior Door C\_East, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_TC_A9}
\caption{Temperature at exterior Door C\_East, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_TC_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_GAS}
\caption{Gas Concentration, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_GAS}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_HF}
\caption{Heat Flux, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_TC_A1}
\caption{Temperature at TC Tree A1, Room A, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_TC_A2}
\caption{Temperature at TC Tree A2, Room A, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_TC_A3}
\caption{Temperature at TC Tree A3, Room B, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_TC_A4}
\caption{Temperature at TC Tree A4, Room C, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_2_TC_A5}
\caption{Temperature at TC Tree A5, Room C, Test FIRE\_FA2S1C1NCX1\_2}
\label{fig:FIRE_FA2S1C1NCX1_2_TC_A5}
\end{figure}

\clearpage

\subsection{FIRE\_FA2S1C1NCX1\_3}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_BDP_A7}
\caption{Velocity at exterior Door A, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_BDP_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_TC_A7}
\caption{Temperature at exterior Door A, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_BDP_A8}
\caption{Velocity at exterior Door C\_West, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_BDP_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_TC_A8}
\caption{Temperature at exterior Door C\_West, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_TC_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_BDP_A9}
\caption{Velocity at exterior Door C\_East, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_BDP_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_TC_A9}
\caption{Temperature at exterior Door C\_East, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_TC_A9}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_GAS}
\caption{Gas Concentration, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_GAS}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_HF}
\caption{Heat Flux, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_TC_A1}
\caption{Temperature at TC Tree A1, Room A, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_TC_A2}
\caption{Temperature at TC Tree A2, Room A, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_TC_A3}
\caption{Temperature at TC Tree A3, Room B, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_TC_A4}
\caption{Temperature at TC Tree A4, Room C, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Script_Figures/FIRE_FA2S1C1NCX1_3_TC_A5}
\caption{Temperature at TC Tree A5, Room C, Test FIRE\_FA2S1C1NCX1\_3}
\label{fig:FIRE_FA2S1C1NCX1_3_TC_A5}
\end{figure}

\clearpage

\chapter{Conclusions}
\label{chap:Conclusions}

\chapter{Acknowledgements}
\label{chap:Acknowledgements}

\bibliography{../../../Bibliography/FDS_refs,../../../Bibliography/FDS_general,references}

\appendix

\chapter{Appendix A}

Placeholder

\end{document}