### 2014-07-21 Studies/2012_Chicago/Report/Chicago_Fire.tex
Chicago: continued update to report write-up

\documentclass[11pt,oneside]{book}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                                                 %
% The mathematical style of these documents follows                                               %
%                                                                                                 %
% A. Thompson and B.N. Taylor. The NIST Guide for the Use of the International System of Units.   %
%    NIST Special Publication 881, 2008.                                                          %
%                                                                                                 %
% http://www.nist.gov/pml/pubs/sp811/index.cfm                                                    %
%                                                                                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{../../../Bibliography/commoncommands}

% Rename chapter headings
\renewcommand{\chaptername}{Section}
\renewcommand{\bibname}{References}

% Math shortcuts
\renewcommand{\sb}[1]{_\mathrm{#1}}
\renewcommand{\C}{\mbox{C}}
\renewcommand{\H}{\mbox{H}}
\renewcommand{\O}{\mbox{O}}
\newcommand{\N}{\mbox{N}}

\begin{document}

\bibliographystyle{unsrt}
\pagestyle{empty}

\begin{minipage}[t][9in][s]{6.25in}

\headerB{
Simulation of an Attic Fire \\
in a Wood Frame Residential Structure \\
Chicago, IL \\
}

\headerC{
\flushright{
Craig G. Weinschenk \\
Daniel Madrzykowski \\
Kristopher J. Overholt \\
{\em Fire Research Division \\
Engineering Laboratory \\
Gaithersburg, Maryland, USA} \\ }
}

\flushright{\today \\
}

\vfill

\flushright{
\includegraphics[width=2.in]{../../../Bibliography/nistident_flright_vec} \\[.3in]
}

\titlesigs

\end{minipage}

\newpage

\frontmatter

\pagestyle{plain}
\pagenumbering{roman}

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{Contents}
\tableofcontents

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{List of Tables}
\listoftables

\chapter{List of Acronyms}

\begin{tabbing}
\hspace{1.5in} \= \\
BC \> Battalion Chief \\
E \> Engine \\
FDS \> Fire Dynamics Simulator \\
HGL \> Hot Gas Layer \\
HRR \> Heat Release Rate \\
HRRPUA \> Heat Release Rate per Unit Area \\
LODD \> Line of Duty Death \\
NIOSH \> National Institute for Occupational Safety and  Health \\
NIST \> National Institute of Standards and Technology \\
SFPE \> Society of Fire Protection Engineers \\
SQ \> Squad \\
TL \> Tower Ladder \\
\end{tabbing}

\mainmatter

\chapter{Introduction}
Part of the mission of the Fire Research Division at the National Institute of Standards and Technology (NIST) is to develop and apply technology, measurements and standards, improve the understanding of the behavior, prevention and control of fires to enhance fire fighting operations and equipment, fire suppression, fire investigations, and disaster response. NIST has previously used the Fire Dynamics Simulator (FDS) to provide insight on the fire development and thermal conditions of other line of duty death (LODD) fires~\cite{Madrzykowski:1,Iowa,Texas,Cook_County,Grosshandler:Station,Bryner:Charleston,barowy:texas}. The overriding objective of all of these studies is to improve the safety of fire fighters.

On November 2, 2012, a fire in a 2 1/2 story wood frame residential structure claimed the life of a Chicago firefighter. NIST examined the fire dynamics of this incident at the request of NIOSH and Chicago Fire Department (CFD). Computer simulations of the fire using Fire Dynamics Simulator and Smokeview, provide insight on the fire development and thermal conditions that may have existed in the residence during the fire. The specific objectives of the simulations detailed in this report are: 
\begin{enumerate}
\item To examine with physics-based calculations the effect of flow path on the temperature, pressure, and fire conditions in this multi-story residential structure.
\item To provide visualization of fire behavior representative of what Chicago Fire Department members experienced during the course of their interior operations.
\end{enumerate}
This document describes the input and the results of the FDS (version 6.0.0) simulations.

\chapter{Fire Incident Summary}
\label{fire_sum}
The following account of events is based on information provided by NIST physical measurements on-scene, NIST photographs of the scene, time line information provided by the Chicago Fire Department and information provided from the NIOSH Fire Fighter Fatality Investigation and Prevention Program Report~\cite{NIOSH:Bowyer}.

On November 2, 2012, a fire in a 2 1/2 story wood frame residential structure resulted in the death of a fire captain and injury to a fire fighter. At approximately 5:16 in the afternoon, a dispatch was made for ``smoke in the area''. Within a minute of dispatch, a Battalion Chief (BC) arrived and sized up the structure fire from the exterior and reported a working fire with heavy smoke coming from the front and rear of the structure's attic. After speaking with the building occupants, the BC entered the structure and found light smoke on the second floor and a glow around the doorway that led from the second floor hallway to enclosed rear porch. Engine 123 (E123) arrived on scene at approximately 17:19. The victims were part of the E123 crew.  The BC met the Captain and the Firefighter (victims) at the front door to the structure.  Then the Captain and Firefighter took their hose line to the second floor of the structure. At approximately 5:21, the Captain and Fire Fighter were on the second floor with their hoseline and had begun advancing the charged line toward the rear second floor door between the apartment and the enclosed porch. Approximately 30 seconds later, a fire fighter from Tower Ladder 39 (TL39) enters the enclosed porch on the rear of the structure by opening the rear exterior door to the first floor of the porch. The firefighter from TL39, notices ``the fire light up'' in the stairwell of the enclosed porch, and then kicks in the first level door between the rear porch and the main structure. Both doors remain open and the firefighter from TL39 exits the structure. Given the amount of fire showing from the second floor of the rear porch and ``heavy fire'' in the stairwell of the rear porch, the BC radios the E123 Captain at 17:23, to alert him to the fire conditions and to advise him that E49 would be applying water to the fire on the rear of the structure from the exterior. No response was made by the E123 Captain. It is believed that the door between the rear porch and the second floor hallway, where the E123 Captain and firefighter were located may have failed by this time.  Approximately a minute later E49 applied water to the fire in the attic.

Squad 5 (S5) entered the front of the structure at approximately 17:25 and headed to the second floor. They heard a commotion and found the E123 Captain collapsed on the kitchen floor and entangled with the E123 firefighter. A Mayday was called on the radio by S5 at approximately  17:27 and the E123 Captain was removed from the front of the building at approximately 17:29 and succumbed to his injuries at the hospital at 17:38. Table~\ref{tab:fire_info} presents the timeline of the fire.

\begin{table}
\centering
\captionof{table}{Fire Event Timeline}\label{tab:fire_info}
\begin{tabular}{l l}
\toprule[1.5pt]
Incident Time [hr:min:s] & Fire Behavior / Fireground Operation  \\
\midrule
\multirow{2}{*}{17:15:00} & \multirow{2}{*}{\parbox{10cm} {Reported working fire with heavy smoke coming from the front and rear of the structure's attic.}} \\ 
         & \\[.25cm] %blank rows exist for wrapping text. # of blanks is 1 less than number of rows
\multirow{2}{*}{17:16:00}    &  \multirow{2}{*}{\parbox{10cm} {Dispatch for a ``Still'' alarm $-$ two engine companies, two truck companies, and a battalion chief.}} \\
         & \\[.25cm]
\multirow{2}{*}{17:20:00}    & \multirow{2}{*}{\parbox{10cm} {Attic self ventilation - heavy smoke out of front attic window and heavy smoke and fire out of rear.}} \\
         & \\[.25cm] 
17:21:00 & E123 Capt. and FF on 2$^{nd}$ floor with 1 3/4 in hose line. \\[.25cm]
17:21:30 & 1$^{st}$ floor exterior porch door opened. \\[.25cm]
\multirow{2}{*}{17:23:00}    & \multirow{2}{*}{\parbox{10cm} {Heavy fire report in rear stairwell and enclosed porch. No response from E123: Capt. or FF.}} \\
         & \\[.25cm]
17:24:00    & Water applied to attic from the rear exterior position. \\[.25cm]
17:25:00    & SQ5 makes entry through front of structure. \\[.25cm]
17:27:00    & After hearing a Mayday call from SQ5, IC calls a Mayday. \\[.25cm]
17:29:00    & E123 Capt. carried out of structure. \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

\chapter{Model Description}
\label{model}
FDS is a computational fluid dynamics (CFD) model that solves a form of the Navier-Stokes equations appropriate for low-speed, thermally driven flow with an emphasis on smoke and heat transport from fires.  Within a CFD model, the room or building is divided into small three-dimensional rectangular control volumes or computational cells.  The cells are contained together within one larger volume known as a computational domain.  The CFD model computes the density, velocity, temperature, pressure and species concentration of the gas in each cell.  Based on the laws of conservation of mass, momentum,  and energy, the model tracks the generation and movement of fire gasses. One of the most important advantages of FDS is that it is  mathematically verified~\cite{FDS_Verification_Guide} and validated against fire test data to ensure that it provides the expected results, given sufficient input data~\cite{FDS_Validation_Guide}.  A complete description of the FDS model is given in reference~\cite{FDS_Math_Guide}.

Smokeview is a software tool designed to visualize the results of the FDS~\cite{Smokeview_Users_Guide}.  Smokeview visualizes smoke and other attributes of the modeled fire using traditional scientific methods such as displaying tracer particle flow, two dimensional (2D) or three dimensional (3D) shaded contours of gas flow data such as temperature and flow vectors showing flow direction and magnitude.  Smokeview allows the fire and fire movement to be visualized.  This is done by displaying a series of partially transparent planes where the transparencies in each plane (at each grid node) are determined from soot densities computed by FDS.  Smokeview also visualizes static data at particular times using 2D or 3D contours of data such as temperature and flow vectors showing flow direction and magnitude.

Input data from a number of sources must be collected and documented in order to simulate a fire with FDS or any other model. For the model presented here, information was obtained from two main sources. From the fire scene the following information was gathered: the geometry of the building and the compartments being modeled, the size and location of exterior and interior ventilation openings, and documentation of fire damage to the building. From witnesses, first responders, reports, and recorded media such as fire ground radios or videos: information on the timing of the fire development, order and approximate timing of ventilation openings to the outside, and weather conditions at the time of the fire. The following subsections will describe the inputs used to construct the simulation, including any engineering approximations or assumptions that were made.

\section{Geometry}
\label{geom}
The structure was a 2 1/2 story residential building with balloon frame construction and gable roof. \footnote{Balloon frame construction is a method of framing where a continuous stud runs from the bottom horizontal member to the top plate. Intermediate floors are built directly off of the studs. One potential issue with balloon frame construction, is that without fire stops, there is void space for fire to potentially travel between the floors.} The structure had a 17.15~m (56~ft) by 6.15~m (20~ft) footprint (cf. Fig.~(\ref{fig:simp_geom})). This results in a floor area of approximately 105~m$^2$ (1100~ft$^2$) per floor and a main living area of approximately 110~m$^2$ (2200~ft$^2$). Interior walls were constructed from nominal lumber with 12.7~mm (0.5~in) gypsum board over top~\cite{NIOSH:Bowyer}. Additionally there was a half-story attic that was used to house the structure's hot water heater. There were three sets of stairs in the structure: one front stairwell that connected the first and second levels (Stair 1 in Fig.~(\ref{fig:simp_geom})), one rear stairwell that connected the first and second levels (Stair 2 in Fig.~(\ref{fig:simp_geom})) and a second rear set of stairs which provided access to the attic (Stair 3 in Fig.~(\ref{fig:simp_geom})). Entry to the first floor of the structure can be made either in the front or rear, see Fig.~(\ref{fig:simp_geom}). The fire, which is described in Section~\ref{fire}, started in the attic and spread to the rear porch. Discussion of the fire's impact on interior conditions can be found in Section~\ref{results}. Fig.~(\ref{fig:alpha_ex}) shows the post-incident exterior of the front and rear of the structure. Note that the majority of the damage is to the structure's attic and second floor rear porch. Fully dimensioned drawings of the first and second floor interiors can be found in Figs.~(\ref{fig:first_floor} and \ref{fig:second_floor}), respectively.

\begin{figure}[h!]
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
\includegraphics[width=\textwidth]{../Figures/1st_Floor_Metric} \\
\includegraphics[width=\textwidth]{../Figures/2nd_Floor_Metric}
\end{tabular*}
\caption{Plan view of 1st (top) and 2nd (bottom) floor of structure. Stairs and key rooms are identified.}
\label{fig:simp_geom}
\end{figure}

\begin{figure}[h!]
\centering
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
   \scalebox{1.0}{ \includegraphics[width=3in]{../Figures/exterior_alpha} } &
   \scalebox{1.0}{ \includegraphics[width=2.67in]{../Figures/exterior_charlie} } \\
\end{tabular*}
\caption{Photograph showing the post-incident front (left) and rear (right) exterior of the structure.}
\label{fig:alpha_ex}
\end{figure}
%Fig.~(\ref{fig:smv_interior}) shows 3D Smokeview renderings of cutaways of the beta and delta sides of the structure interior.
%\begin{figure}[h!]
%\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
%\includegraphics[width=.80\textwidth]{../Figures/interior_beta} \\
%\includegraphics[width=.80\textwidth]{../Figures/interior_delta} \\
%\end{tabular*}
%\caption{3D Smokeview rendering of structure interior looking from the charlie side to both the beta side (top) and to the delta side (bottom)}
%\label{fig:smv_interior}
%\end{figure}

\section{Fire}
\label{fire}

To estimate the fuel, fire size, and structure ventilation, on-scene videos and post-incident reports are used. Based on these findings, the FDS model includes two source fire locations: an initial fire in the attic and a secondary fire on the second floor porch landing. The attic and rear porch consisted of a fuel load that is mainly wood: the rear porch was enclosed with wood planks exteriorly covered by foam board and vinyl siding, the roof was asphalt shingles over wood planking, and the attic and porch had oriented strand board (OSB) floors.

Wood can be represented by the the following chemical formula, $\C\H_{1.7}\O_{0.74}\N_{0.002}$, with yields of carbon monoxide ($y_{\mathrm{CO}}=0.004 \; {\mathrm{kg_{CO}}/\mathrm{kg_{wood}}}$) and soot ($y_{\mathrm{C}}=0.015 \; {\mathrm{kg_{C}}/\mathrm{kg_{wood}}}$) found from the SFPE Handbook~\cite{SFPE:Tewarson}. A balanced chemical reaction for wood oxidation can written:

\begin{multline}
\C\H_{1.7}\O_{0.74}\N_{0.002} + 4.89(0.208\,\O_{2} + 0.783\,\N_{2} + 0.387\text{\sc{e}-}3\,\C\O_{2} + 0.834\text{\sc{e}-}2\,\H_{2}\O) \\ 
\rightarrow 5.84(0.719\text{\sc{e}-}3\,\C\O + 0.168\,\C\O_{2} + 0.155\,\H_{2}\O + 0.669\,\N_{2} + 0.693\text{\sc{e}-}2\,\C)
\label{eq:wood_comb}
\end{multline}

The heat of combustion of wood used in this simulation is 16400 kJ/kg, based on data provided in the SFPE Handbook~\cite{SFPE:Tewarson}. The heat of combustion quantifies the energy release per unit mass of the fuel. The following text defines Eq.~\ref{eq:wood_comb} for an FDS input file with the desired fuel and reaction properties discussed above:

\begin{lstlisting}
&SPEC ID = 'WOOD', FORMULA = 'CH1.7O0.74N0.002' /

&REAC ID = 'wood' 
    FUEL = 'WOOD', 
    HEAT_OF_COMBUSTION=16400,
    SOOT_YIELD = 0.015,
    CO_YIELD = 0.004/
\end{lstlisting}

Note that using the input lines above will invoke the simple chemistry reaction balance in FDS. This means that fuel and air react to form only CO$_2$, CO, H$_2$O, Soot, and N$_2$. If there is a desire for any other combustion products then the user must explicitly define those species and the chemical reaction that produces them~\cite{FDS_Users_Guide}. Based on above input lines FDS uses the default, mixing-controlled fast chemistry combustion model. This mechanism  states that the rate of fuel consumption is proportional to both the local limiting reactant concentration and the local rate of mixing and extinction is based on the critical flame temperature approach~\cite{FDS_Math_Guide}. While FDS provides users to the option to use a more complex finite-rate combustion mechanism, we do not have sufficient evidence to justify deviating from the default specifications.

To estimate the energy release per unit time and unit area (kW/m$^2$) of wood, we use work by Babrauskas and Grayson~\cite{babrauskas1990}. Experiments were conducted in a cone calorimeter \footnote{The cone calorimeter is an experimental apparatus used to gather data about ignition time, mass loss, combustion products, and heat release rate among other properties associated with burning small samples of materials.} to determine the 5 minute average of HRRPUA  \footnote{HRRPUA is the heat release rate per unit area.} (kW/m$^2$) for several different types of wood over a range of radiant heat fluxes (kW/m$^2$). Babrauskas and Grayson~\cite{babrauskas1990} suggest that the heat release rate per unit area for a variety of woods is approximately 50~kW/m$^2$.

To calculate the total fire size, the appropriate burnable area (exposed wood area) of the attic and second floor porch enclosure needs to be determined. Based on the floor plan of the structure, the underside of the roof is approximately 134~m$^2$ (1453~ft$^2$) with 7~m$^2$ (75~ft$^2$) walls on the front and rear sides. Additionally there is 105~m$^2$ (1130~ft$^2$) of floor area in the attic. Examination of post incident images showed that the floor immediately over the back porch burned away. That area will be considered as part of the porch fire. Based on post-incident damage, approximately 50 \% of the remaining floor area as fuel to the attic fire and 90 \% of the roof is con sidered. To estimate the total attic fire size, the underside roof area (120~m$^2$), two end walls (14~m$^2$) and the floor area (43~m$^2$ (463~ft$^2$)) are included. Using an HRRPUA of wood of 50~kW/m$^2$, results in a potential peak fire size of 9~MW. To estimate the porch fire size, the the four walls of porch enclosure (42~m$^2$) and the floor and ceiling of the porch (24~m$^2$ (258 ft$^2$)) are included. The total area is 64~m$^2$ (689~ft$^2$). Using the same HRRPUA of 50~kW/m$^2$, the potential peak fire size on the porch is approximately 3.3~MW. The total fire size for this structure is estimated to be at peak, 12.3~MW. Since it is unknown when the fire started relative to notification of a fire or how fast the fire spread, several simplifications are made. First is that the attic fire starts at the beginning of the simulation and ramps up to 9~MW in 10 seconds. The porch fire starts 40 seconds into the simulation and ramps to 3.3~MW in 10 seconds. A ten second ramp to peak was used to prevent sharp velocity changes at the fire boundary condition in FDS. The second is that the sources of the gas phase wood fuel are of fixed area and at fixed locations (2 sources for attic, 1 source for porch). The following lines define the source fires for this FDS input file:

\begin{lstlisting}
&SURF ID='FIRE', HRRPUA=2250., COLOR='ORANGE',TAU_Q=10 /
&VENT XB=-13.5,-13,1,5,6.16,6.16, SURF_ID='FIRE' / 2 m^2
&VENT XB=-14.5,-14,1,5,6.16,6.16, SURF_ID='FIRE' / 2 m^2
&VENT XB=-16.5,-15.5,1,2.5,3.14,3.14, SURF_ID='FIRE',DEVC_ID='SECOND FIRE'/ 1.5 m^2
&DEVC XYZ = -17.2,2,0 ID = 'SECOND FIRE', QUANTITY = 'TIME', SETPOINT = 40, INITIAL_STATE = .FALSE. /
\end{lstlisting}
Note that the coordinates here are unique for the vents are specific to the input files used for these simulations. The key point is that they are 2-D planes of a specific area.

\section{Materials}
\label{matl}
While the fire was simplified to constant area fuel sources (Section~\ref{fire}), defining the material properties of the walls and ceiling are still important from a heat transfer perspective. In this study, gypsum~\cite{Drysdale:1} is used for the finished walls and ceilings on the main two floors of the structure and wood~\cite{Incropera:1} for the unfinished portions of the attic and rear porch. The text provided below are the lines necessary to define these materials in this FDS input file.

\begin{lstlisting}
&MATL ID        = 'GYPSUM'
    FYI = 'Drysdale' 
    CONDUCTIVITY    = 0.17
    SPECIFIC_HEAT   = 1.09
    DENSITY = 930./

&MATL ID        = 'WOOD_MAT'
    FYI = 'Incropera and DeWitt, Fundamentals of Heat and Mass Transfer'
    CONDUCTIVITY    = 0.12
    SPECIFIC_HEAT   = 1.38
    DENSITY = 510./ 
\end{lstlisting}

\section{Ventilation}
\label{Vents}
The simulations in this study account for the changes in ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). There are two ventilation scenarios that are considered in this study. The first is the baseline simulation where the ventilation times and ventilation areas represent our best understanding of the incident. The alternative simulation does not include the firefighter opening the rear door to make entry to the structure and the subsequent exterior left wall fire penetration. The times of the ventilation changes are provided in Table \ref{tab:vents}. Structure leakage has been shown to be important when modeling enclosure fires~\cite{beal2009}. The open front door at the start of the simulation and the vents (opening size and open time) created during the simulation are significantly larger than the total effective area of leakage of a structure of this type. Therefore, leakage would have a negligible impact on the fluid mechanics within the structure and is not included in this study.

\begin{table}
\centering
\captionof{table}{Simulation Ventilation Timelines}\label{tab:vents}
\begin{tabular}{ccccc}
\toprule[1.5pt]
Baseline Simulation [s] & Alternative Simulation [s] & Ventilation Area [m$^2$] & Type & Structure Side \\
\midrule
0    &  0   & 0.11 & Window      & Rear \\
40    &  40   & 0.08 & Breach (x2) & Front and Rear \\
45    &  45   & 0.12 & Breach      & Left \\
55    &  55   & 0.12 & Breach      & Rear \\     
130   &  N/A  & 1.41 & Door        & Rear \\
135   &  N/A  & 0.2  & Breach      & Left \\        
160   &  160  & 0.80 & Door        & Interior \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

\section{Mesh}
\label{mesh}

For the simulation in this study, we can estimate a measure of how well the flow field is resolved by the non-dimensional expression $D^*/\dx$. Here $D^*$ is a characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g} }
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}   
From the FDS User Guide~\cite{FDS_Users_Guide}, ``the characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$.'' Here, $D$ is the physical diameter of the fire as input to the model. Following Eq.~(\ref{eq:mesh}) and using 10 cm grid cells, the characteristic fire diameter to cell size ratio based on the 3.3 MW porch fire is 15.5 and for the 9 MW fire is 23. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16~\cite{NUREG_1824}. The grid resolution used in this model falls within typical, engineering $D^*/\dx$ values. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10 cm. Fig.~(\ref{fig:geom_grid}) shows the 
front of the structure rendered in Smokeview with the 10 cm mesh.

\begin{figure}[h!]
\centering
\includegraphics[width=.60\textwidth]{../Figures/smv_exterior_grid}
\caption{Front side of the structure with 10 cm computation mesh.}
\label{fig:geom_grid}
\end{figure}

As discussed Section~\ref{geom}, the structure has 17.15~m (56~ft) by 6.15~m (20~ft) footprint with a total height of 8.56~m (28~ft). To ensure adequately resolved fluid flow into and out of the structure the computational domain is extended beyond the volume of the structure. The total volume of the computational domain is 24~m (79~ft) by 10~m (33~ft) by 10~m (33~ft). At 10~cm (3.9~in) resolution this would result in 2.4 million computational cells. To make the simulation time more tractable, the domain was divided into 16 equal 150 thousand cell meshes to be processed in parallel. Fig.~(\ref{fig:mult_mesh}) shows the entire structure within the computational domain.

\begin{figure}[h!]
\centering
\includegraphics[width=.60\textwidth]{../Figures/smv_exterior_mesh}
\caption{Front-right side of the structure within computational domain. Each of the 16 meshes is 3 m (9.8 ft) by 10 m (33 ft) by 5 m (16.4 ft).}
\label{fig:mult_mesh}
\end{figure}

For the simulations performed for this study, Table~\ref{tab:mod_param} includes the relevant model input parameters discussed in Section~\ref{model}.

\begin{table}
\centering
\captionof{table}{Relevant Modeling Input Parameters}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter & Description & Reference  \\
\midrule
 Simulation Time           & 5 minutes                             & -- \\[.25cm]
 Grid Size                 & 10 cm                                 & Section~\ref{mesh} \\ [.25cm]
 Ambient Temperature       & 25 $^{\circ}$C                        & -- \\ [.1cm]
\multirow{4}{*}{Reaction: Wood}  & Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$              & \multirow{4}{*}{Section~\ref{fire}} \\
                           & CO Yield: 0.004 \; ${\mathrm{kg_{CO}}/\mathrm{kg_{wood}}}$  & \\
                           & Soot Yield: 0.015 \; ${\mathrm{kg_{C}}/\mathrm{kg_{wood}}}$ & \\
                           & Heat of Combustion: 16400~kJ/kg                             & \\ [.25cm]
\multirow{2}{*}{Peak HRR}  & Attic Fire: 9~MW                                            & \multirow{2}{*}{Section~\ref{fire}}\\ 
                           & Porch Fire: 3.3~MW                                          & \\ [.25cm]                     
\multirow{3}{*}{Material: Wood}    & k: 0.12 \; \si{W/(m.K)})        & \multirow{3}{*}{Section~\ref{matl}} \\
                                   & $\rho$: 510 \; kg/m$^3$         & \\
                                   & c$_{p}$: 1.38 \; \si{kJ/(kg.K)} & \\[.25cm]
\multirow{3}{*}{Material: Gypsum}  & k: 0.17 \; \si{W/(m.K)}         &  \multirow{3}{*}{Section~\ref{matl}}\\ 
                                   & $\rho$: 930 \; kg/m$^3$        & \\ 
                                   & c$_{p}$: 1.09 \; \si{kJ/(kg.K)} & \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

\section{Timeline}
\label{timeline}
From the NIOSH report~\cite{NIOSH:Bowyer}, a timeline of events was able to be created, see Table~\ref{tab:fire_info}. Due to a lack of information about when the fire ignited, an ignition time/start of the simulation needed to be estimated.  We estimate that the simulation starts at 17:19:20. Since the actually time was unknown, the fire growth rate in the computational model is such that when firefighters enter the structure, the hazard level in the model is similar to the real situation. A second unknown that was estimated was the time when the second floor door fails (2:40 into the simulation). While this time is not exact, the takeaway from the analysis in the following sections should be impact of the door's failure on the interior conditions of the structure. Table~\ref{tab:firesim} shows the fireground timeline along with the corresponding simulation times.

\begin{table}
\centering
\captionof{table}{Fire - Simulation Event Timeline}\label{tab:firesim}
\begin{tabular}{l l l}
\toprule[1.5pt]
Incident Time [hr:min:s] & Simulation Time [s] & Fire Behavior / Fireground Operation  \\
\midrule
\multirow{2}{*}{17:15:00} &    & \multirow{2}{*}{\parbox{10cm} {Reported working fire with heavy smoke coming from the front and rear of the structure's attic.}} \\ 
         & & \\[.25cm] %blank rows exist for wrapping text. # of blanks is 1 less than number of rows
\multirow{2}{*}{17:16:00} & \multirow{2}{*}{}   &  \multirow{2}{*}{\parbox{10cm} {Dispatch for a ``Still'' alarm $-$ two engine companies, two truck companies, and a battalion chief.}} \\
         & \\[.25cm]
         &  \multirow{2}{*} {0} &  \multirow{2}{*}{\parbox{10cm} {Simulation starts. Attic fire ignites. Initial self ventilation on charlie side. Front door open.}} \\
         & & \\[0.25cm]
\multirow{2}{*}{17:20:00} & \multirow{2}{*}{40}  & \multirow{2}{*}{\parbox{10cm} {Attic self ventilation - heavy smoke out of front attic window and heavy smoke and fire out of rear. Rear porch ignites in model.}} \\
         & & \\[.25cm] 
         & 45 & Model ventilation on charlie side. \\[.25cm] 
         & 55 & Model ventilation on beta side. \\[.25cm] 
17:21:00 & 100 & E123 Capt. and FF on 2$^{nd}$ floor with 1 3/4-in hose line. \\[.25cm]
17:21:30 & 130 & 1$^{st}$ floor exterior porch door opened. Attic fire ignites in model. \\[.25cm]
        & 135 & Additional model ventilation on beta side. \\[.25cm] 
        & 160 & Top of interior door fails in model. \\[.25cm] 
\multirow{2}{*}{17:23:00}    & \multirow{2}{*}{220} & \multirow{2}{*}{\parbox{10cm} {Heavy fire report in rear stairwell and enclosed porch. No response from E123: Capt. or FF.}} \\
         & & \\[.25cm]
17:24:00    &   & Water applied to attic from the rear exterior position. \\[.25cm]
17:25:00    &   & SQ5 makes entry through front of structure. \\[.25cm]
\multirow{2}{*}{17:27:00}    & \multirow{2}{*}{300}  & \multirow{2}{*}{\parbox{10cm} {After hearing a Mayday call from SQ5, IC calls a Mayday. Simulation ends.}}\\
         & &  \\[.25cm]
17:29:00    &   & E123 Capt. carried out of structure. \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

\chapter{Model Results}
\label{results}

\section{Heat Release Rate}
\label{HRR}
The heat release rate output data from FDS quantifies the simulated rate of energy release in the simulation. Fig.~(\ref{fig:hrr}) shows the HRR versus time based on the prescribed fires described in Section \ref{fire} and the FDS results for two venting scenarios. The two vertical lines represent when a firefighter opened the back door of the structure (t = 130 s) and when the second floor interior door was estimated to fail (t = 160 s). Due to a lack of information about when the fire ignited or how fast it grew, these parameters needed to be estimated.  While the growth rate shown in Fig.~(\ref{fig:hrr}) (10~MW in approximately 20 seconds) is atypically fast, analysis of the incipient stage of the fire growth was not considered in this study. The goal was to have a fire approach its estimated peak and become ventilation limited quickly to minimize computation time while aligning with the constraints of videos taken during the fire. Accuracy of heat transfer to the walls of the structure may be sacrificed by taking this approach, however all window, door, and wall failure were driven by the event timeline not modeling predicted failure temperatures.

\begin{figure}[h!]
\centering
\includegraphics[width=.80\textwidth]{../Figures/Chicago_Fire_HRR}
\caption{HRR versus time of FDS model of the West 50th St. fire. The first vertical line indicates when the back door failed, the second indicates when the interior hallway door failed.}
\label{fig:hrr}
\end{figure}

In Fig.~(\ref{fig:hrr}), the solid line is the baseline FDS simulation (this represents the on-scene conditions). The dashed line represents an alternative series events, specifically the rear door does not get opened and the subsequent beta side breach does not occur. The dash-dotted line is the prescribed FDS input HRR. We can see that within the first 10 seconds, the structure is under-ventilated as the modeled HRR is below the prescribed HRR. The HRR peaks at just under 9 MW before becoming oxygen limited. The HRR drops to 5~MW as the available oxygen inside the structure gets consumed. There is insufficient available oxygen in the structure to combust all of the fuel and combustion occurs locally at vents. Fig.~(\ref{fig:smv_ext_fire}) illustrates exterior flaming combustion through the failed Charlie side window and attic breach just after the second fire ignites at 40 seconds. After additional wall breaches at 45 and 55 seconds (see Table~\ref{tab:vents}), the abrupt change in ventilation causes the HRR spikes to 9~MW before again dropping to about 6~MW. The HRR then begins to increase towards a sustainable steady state value.

From Fig.~(\ref{fig:hrr}), we see that the HRR for both FDS simulations are identical until 130 seconds (HRR $\approx$ 7~MW). At 130 seconds the rear door is opened in the baseline simulation. At 135 seconds there is an additional beta side breach (Table~\ref{tab:vents}). The open door and breach act as additional pressure releases for the structure. Again, the abrupt change in ventilation causes the HRR to spike. Here we see a rise from 7~MW to 9.7~MW at 139 seconds. At the time of door failure (160 seconds), the HRR in both cases was approximately 9 MW as there was sufficient ventilation. While there might be some locally different burning characteristics between the baseline simulation and the alternate simulation, in both cases the interior conditions represented a significant hazard to firefighters inside the structure. Therefore, the analysis will focus on the results of the baseline simulation.

\begin{figure}[h!]
\centering
\includegraphics[width=.675\textwidth]{../Figures/smv_exterior_fire}
\caption{Smokeview rendering showing combustion occurring exterior to the structure on the Charlie side.}
\label{fig:smv_ext_fire}
\end{figure}

\section{Pressure}
As the fire grows in the attic and back porch, the pressure rises. Fig.~(\ref{fig:pres_159s}) shows the pressure rise in the structure just before the interior door failed in the simulation. On the rear porch, the pressure rise ranges from a 4~Pa over pressure 1~m off the floor to greater than 10~Pa at the ceiling.
\begin{figure}[!ht]
\includegraphics[width=.75\textwidth]{../Figures/west_50th_baseline_pres2}
\input{../Scripts/colorbar_metric_pres}
\caption{FDS simulated pressure contours, 1 second prior to interior door failure. Note 1~Pa is 1.4 $\times 10 ^{-4}$ psi.}
\label{fig:pres_159s}
\end{figure}
Gases at elevated pressure will always travel toward an area of lower pressure. The door acts as a boundary between the high and low pressure areas. Once the door fails open, the high pressure gases flow toward the low pressure areas in the house, hallway, kitchen, and front stairwell. The velocity vectors in Fig.~(\ref{fig:velo_175s}), 15 seconds after the interior door fails, confirms this flow direction within the structure.
\begin{figure}[!ht]
\includegraphics[width=.675\textwidth]{../Figures/west_50th_baseline_velo_175}
\input{../Scripts/colorbar_std_velo}
\input{../Scripts/colorbar_metric_velo}
\caption{FDS simulated velocity vectors, 15 second after interior door failure.}
\label{fig:velo_175s}
\end{figure}

\section{Temperature}
\label{temp}
Analysis of the simulation temperatures will focus on the high hazard areas of the structure: the attic, the rear porch and stairwell, and the interior hallway which connects the rear porch to the front stairwell. Here, we will examine the temperature profile through these high hazards area at several snapshots in time to illustrate the changing interior conditions. The first times are 129 seconds into the simulation, just before a firefighter opened the rear porch door (cf. Fig.~(\ref{fig:temp_129s})) and 159 seconds into the simulation, just before the interior door fails (cf. Fig.~(\ref{fig:temp_159s})).

\begin{figure}[!ht]
\includegraphics[width=.675\textwidth]{../Figures/west_50th_baseline_129}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{FDS simulated temperature contours, 1 second prior to rear door opening.}
\label{fig:temp_129s}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=.675\textwidth]{../Figures/west_50th_baseline_159}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{FDS simulated temperature contours, 1 second prior to interior door failure.}
\label{fig:temp_159s}
\end{figure}

Fig.~(\ref{fig:temp_129s}) shows temperatures of at least 260~$^{\circ}$C (500~$^{\circ}$F) in the second floor rear porch and rear portion of the attic. High temperatures, between 80~$^{\circ}$C (185~$^{\circ}$F) and 200~$^{\circ}$C (390~$^{\circ}$F), also extend down to the 1$^{st}$ floor rear entrance. Before the interior door fails, Fig.~(\ref{fig:temp_159s}), the inflow ambient air from the open door lowers the temperature local to the doorway, but we know from Fig.~(\ref{fig:hrr}) that over this interval the HRR spikes because of the additional ventilation. After the second floor door fails, (cf. Fig.~(\ref{fig:temp_161s})), the high temperature gas that filled the back porch flow in the direction of lower pressure (towards the front door). Note that the conditions in the attic do not appear to be influenced.
\begin{figure}[!ht]
\includegraphics[width=.675\textwidth]{../Figures/west_50th_baseline_161}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{FDS simulated temperature contours, 1 second after interior door fails.}
\label{fig:temp_161s}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=.675\textwidth]{../Figures/west_50th_baseline_175}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{FDS simulated temperature contours, 15 seconds after interior door fails.}
\label{fig:temp_175s}
\end{figure}

From Fig.~(\ref{fig:temp_175s}), 15 seconds after the door failed, high temperature gases have descended to approximately 1~m (3.28~ft) off the ground through the interior hallway and have spilled into the kitchen (recall layout from Fig.~(\ref{fig:simp_geom})).

\chapter{Discussion}
The discussion will be broken up into two parts. The first part will address the results of the computational modeling as they relate to the interior hallways conditions. The second part will examine the the model results and observed outcomes of fire as they relate to tactics discussed in recent experimental research.

\section{Hallway Flow Path}
From on-scene reports from the BC (Section~\ref{fire_sum}) and simulation results (Section~\ref{temp}), interior conditions on the second floor of the structure were initially tenable. From the NIOSH report, while the BC was inside the structure on the second floor, the  BC observed only a light haze and a glow around the door between the hallway and enclosed porch. This indicated that some amount of the fire was burring in the area of the enclosed porch. The simulation shows a high hazard area in the attic and the enclosed porch at this time. By the time the E123 Captain and FF arrived on the second floor (Table~\ref{tab:firesim}) and approached the rear of the structure, the rear porch was fully involved. Temperature and pressure, greater than 260~$^{\circ}$C (500~$^{\circ}$F) and 5~Pa respectively, were on the rear side of the steel door at the rear of the structure. Exposure to these conditions weakened the structural integrity of the door, eventually leading to a partial failure. Fig.~(\ref{fig:chicago_doorfold}) shows the state of the rear door post fire. Note the large crease in the door just above the knob.

\begin{figure}[h!]
\centering
\includegraphics[width=.4\textwidth]{../Figures/Porch_Door_1}
\caption{Post fire image of interior door.}
\label{fig:chicago_doorfold}
\end{figure}
Recent experiments conducting by the National Institute of Standards and Technology revealed similar post-fire states after doors were exposed to fires. Fig.~(\ref{fig:SC_doorfold}) shows doors from two different house fire experiments in different stages of degradation. In both cases note that the door knob assembly served as a pivot point for the upper portion of the steel from the door to collapse around. This is similar to the location of the folds in the remains of the steel door that was located between the hallway and the enclosed porch.

\begin{figure}[h!]
\centering
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
   \scalebox{1.0}{ \includegraphics[width=3in]{../Figures/SC_Door_1} } &
   \scalebox{1.0}{ \includegraphics[width=2in]{../Figures/SC_Door_2} } \\
\end{tabular*}
\caption{Post experiment images of door failures.}
\label{fig:SC_doorfold}
\end{figure}

While we have estimated the time of failure for the door, the key take away is that after the door failed an interior flow path from the enclosed porch to the kitchen of the structure was created. This resulted in a rapid change in the conditions within the flow path. The combustion-gases at elevated temperature and pressure  in the enclosed porch were vented by the failure of the door to lower pressure areas in the hallways and kitchen, where the conditions seems to equalize. The arrows in Fig.~(\ref{fig:flowpath_1}) show the flow direction of gases on the second floor following the door failure.

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/ChicagoFlow}
\caption{Flow path on second floor following interior door failure.}
\label{fig:flowpath_1}
\end{figure}

\subsection{Assessing the Hazard}

A person is susceptible to second degree burn injuries when exposed to temperatures greater than 55~$^{\circ}$C (130~$^{\circ}$F) \cite{designation2003c}. While firefighters wear protective gear, gear only offers finite levels of protection. The polycarbonate in a self contained breathing apparatus begins to soften when the material temperature reaches approximately 140~$^{\circ}$C (284~$^{\circ}$F) \cite{mensch2011emergency}. Structural fire fighting coats and pants are tested to withstand temperatures of 260~$^{\circ}$C (500~$^{\circ}$F) \cite{nfpa2013standard}. Extended exposure to elevated temperatures can result in significant heat transferred to the firefighter, putting him/her at risk. Therefore, for analysis of the modeled exposure temperatures, we set the maximum visualized temperature to be all values that are 260~$^{\circ}$C (500~$^{\circ}$F) or above. This means that temperatures visualized in the highest range in the model output are at a minimum of 260~$^{\circ}$C (500~$^{\circ}$F); predicted temperatures could in fact be higher.

Prior to door failure, the elevated temperatures on the second floor were limited to the rear porch, Fig.~(\ref{fig:temp_top_159}). Temperatures are highest at the back-left corner where there is sufficient oxygen due to the section of porch wall that had burned away allowing additional oxygen to mix with the fuel rich fire gases. Five seconds after the door fails, hot gases have traveled the length of the interior hallway and approached the doorway into the kitchen, approximately 4.5~m (14.8~ft) away from the source fire (cf. Fig.~(\ref{fig:temp_top_165})). Within one minute after the door fails, the hallway temperatures are nominally at or above 260~$^{\circ}$C (500~$^{\circ}$F) through out the hallway (cf Fig.~(\ref{fig:temp_top_220})). This illustrates that the high temperature gases seen five seconds after failure were not just a burst but rather the front of a sustained exhaust gas flow from the fire burning in the enclosed rear porch. The conditions in this hallway changed from tenable to high-hazard very rapidly following the door failure. Fig.~(\ref{fig:temp_top_220}) also shows that the temperatures in the kitchen were at or below 90~$^{\circ}$C (194~$^{\circ}$F) and temperatures at the top of the stairs were at or below 70~$^{\circ}$C (158~$^{\circ}$F). At these points in the structure, the exhaust gases in the flow path are air cooled/diluted by the inflow ambient gas. These modeled temperatures are consistent with the post-incident conditions documented in the kitchen and second floor stairwell.
\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/west_50th_baseline_top_159_6ft}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{Temperature contours on the second floor, 1 second prior to door failure, 1.83~m (6~ft) above the floor.}
\label{fig:temp_top_159}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/west_50th_baseline_top_165_6ft}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{Temperature contours on the second floor, 5 seconds after door failure, 1.83~m (6~ft) above the floor.}
\label{fig:temp_top_165}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/west_50th_baseline_top_220_6ft}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{Temperature contours on the second floor, 60 seconds after door failure, 1.83~m (6~ft) above the floor.}
\label{fig:temp_top_220}
\end{figure}

\section{Tactical Considerations}
There have been many fires where the change in a flow path has had an adverse impact on firefighter and occupant safety. Table~\ref{tab:lodd} lists the NIOSH investigation reports from the past 15 years, where it could be determined that it was a flow path related incident. The table lists the NIOSH investigation number, the outcome, and a brief description of the flow path.

Based on a review of these incidents, it is clear that fires with rapidly developing or changing flow paths are a significant hazard to the fire service. The development of, or changes in a flow path could be caused by the failure of a component of the structure, such as a door, window or portion of a ceiling, wall or floor. Environmental conditions such as wind can generate hazardous thermal conditions within a flow path. Uncoordinated ventilation can also be the cause of increased thermal hazards within a flow path.

In this case, the thermal degradation of the steel faced, wood framed, door led to a rapid change in thermal conditions in the hallway resulting in a lethal exposure to victim.  Prediction of the failure mode and failure time of the door is not a reasonable expectation given the large number of unknowns the fire fighters faced on the fire ground.  In this case, the timing of the interior attack appears to have coincided with the failure time of the door. Research has been conducted \cite{madrzykowski2009fire, kerber2009fire} and continues to be conducted by NIST, UL and others that demonstrates that applying water from the exterior into the fire area of a structure, prior to the start of interior operations, can significantly improve the safety of fire fighters by reducing the thermal hazard from the fire and reducing the potential for developing high velocity hot gas flows within the structure.

\begin{table}
\centering
\captionof{table}{Flowpath LODD/LODI Incidents}\label{tab:lodd}
\begin{tabular}{l l l}
\toprule[1.5pt]
NIOSH Report $\#$ & $\#$ LODD/LODI & Flow Path Details  \\
\midrule
\multirow{2}{*}{99-F01 \cite{NIOSH:Pettit} }    & \multirow{2}{*}{3 LODD}                                     & \multirow{2}{*}{\parbox{8cm} {From apartment into hallway on 10th floor of high-rise apartment building}} \\
                 &                                                    & \\
99-F21 \cite{NIOSH:Washenitz}         & 2 LODD/2 LODI                          & Basement to 1st floor \\
F2000-04 \cite{NIOSH:Mezzanotte}   & 3 LODD/3 civilian deaths            & 1st floor to 2nd floor \\
F2000-16 \cite{NIOSH:McFall} & 1 LODD/1 civilian death/1 LODI & 2nd floor hallway through 2nd floor apartment \\
\multirow{2}{*}{F2000-23 \cite{NIOSH:McFall2}} & \multirow{2}{*}{1 LODD/2 LODI       }                   & \multirow{2}{*}{\parbox{8cm} {From ground level to 1st floor then to 2nd floor, flow exited through ceiling}} \\
                 &                                                    & \\
F2000-43 \cite{NIOSH:McFall3}         & 1 serious LODI/2 other LODI      & 1st floor to 2nd floor \\
F2004-02  \cite{NIOSH:Berardinelli}  & 1 LODD                                       & 1st floor to basement \\
F2005-02 \cite{NIOSH:Koedam}        & 1 LODD/4 LODI                           & Rear to front of the building \\
F2005-04 \cite{NIOSH:McFall4}         & 1 LODD/9 LODI                           & Basement to 1st floor \\
F2007-09 \cite{NIOSH:Tarley}           & 1 LODD/2 LODI                           & 3 story training burn $-$ flow through all levels \\
F2007-35 \cite{NIOSH:Braddee}       & 4 LODI                                         & 1st floor to 2nd floor \\
F2009-11 \cite{NIOSH:Merinar}         & 2 LODD                                        & Rear to front of the building \\
F2011-13 \cite{NIOSH:Bowyer2}       & 2 LODD                                        & Lower level up stairs and through entry door and garage \\
F2011-31 \cite{NIOSH:Loflin}            & 1 LODD                                        & Fire extended from lower level apartment \\
F2012-28 \cite{NIOSH:Bowyer}        & 1 LODD/1LODI                            & Attic fire extended into closed porch and then into 2nd floor \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

\chapter{Summary}
NIST's Fire Dynamics Simulator was used to provide insight into the fire dynamics within a 2 1/2 story single family structure in Chicago, Il that resulted in the death of a firefighter. The modeled fuel, fire size, and fire growth rate were estimated taking into account all available information: NIOSH post-incident report~\cite{NIOSH:Bowyer}, post-incident pictures, videos, and relevant literature. This resulted in a specified source fire of 12.3~MW, however based on the limited ventilation in the attic and enclosed porch the model generated a peak HRR of approximately 10 MW as a result of the limited supply of oxygen.  The fire began in the attic and spread down into the enclosed porch on the rear (Charlie side) of the structure.  The spread of heat and fire gases into the second floor of the structure was limited by a closed steel faced, wood framed door that was located between the enclosed porch and the hallway.  After exposure to the elevated temperature and pressure from the fire in the porch area, the wood frame of the door decomposed and the steel faces of the door failed and collapsed inward.  The door failure resulted in the creation of a flow path between the higher pressure and higher temperature conditions in the enclosed porch and the lower pressure and lower temperature conditions in the hallway and kitchen areas of the second floor.  The temperature of the gases that flowed into the hallway at face height exceeded 260~$^{\circ}$C (500~$^{\circ}$F).

Unfortunately a FF and a Captain were advancing a hose line in the hallway at the time of the failure of the door.  The Captain was overcome by the rapid change in conditions before he could exit to the safety of the kitchen area.  After a Mayday, the Captain was rescued from the structure but later succumbed to his injuries.

\chapter{Acknowledgements}
The authors would like to thank the Chicago Fire Department, especially Commissioner Jose Santiago, District Chief Donald Hroma, and District Chief Kevin Krasneck. The authors would also like to thank John Culbertson of Central Valley Fire District, MT for his assistance in compiling Table~\ref{tab:lodd}. Finally, the authors would also like to thank Keith Stakes of NIST and Joseph Willi of University of Illinois Urbana-Champagne for their assistance in creating the drawings of the structure.

\bibliography{../../../Bibliography/FDS_refs,../../../Bibliography/FDS_general,chicago}

\appendix

\chapter{Dimensioned Drawings}

\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{../Figures/50th_Place_1st_Floor}
\caption {Dimensioned drawing of 1$^{st}$ floor. The measurements are accurate to within 15 cm (6 in).}
\label{fig:first_floor}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{../Figures/50th_Place_2nd_Floor}
\caption {Dimensioned drawing of 2$^{nd}$ floor. The measurements are accurate to within 15 cm (6 in).}
\label{fig:second_floor}
\end{figure}

\end{document}