### 2014-07-28 Projects/Spartanburg_2014/Report/Spartanburg_2014.tex
Spartanburg 2014: Add appropriate plots to appendix per test; update test times

\documentclass[12pt,oneside]{book}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                                                 %
% The mathematical style of these documents follows                                               %
%                                                                                                 %
% A. Thompson and B.N. Taylor. The NIST Guide for the Use of the International System of Units.   %
%    NIST Special Publication 881, 2008.                                                          %
%                                                                                                 %
% http://www.nist.gov/pml/pubs/sp811/index.cfm                                                    %
%                                                                                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{../../../Bibliography/commoncommands}

% Rename chapter headings
\renewcommand{\chaptername}{Section}
\renewcommand{\bibname}{References}

% Math shortcuts
\renewcommand{\sb}[1]{_\mathrm{#1}}
\renewcommand{\C}{\mbox{C}}
\renewcommand{\H}{\mbox{H}}
\renewcommand{\O}{\mbox{O}}
\newcommand{\N}{\mbox{N}}

% Center all figures
\makeatletter
\g@addto@macro\@floatboxreset\centering
\makeatother

\begin{document}

\bibliographystyle{unsrt}
\pagestyle{empty}

\begin{minipage}[t][9in][s]{6.25in}

\headerB{
Spartanburg 2014
}

\headerC{
\flushright{
Daniel Madrzykowski \\
Craig G. Weinschenk \\
Kristopher J. Overholt \\
{\em Fire Research Division \\
Engineering Laboratory \\
Gaithersburg, Maryland, USA} \\ }
}

\flushright{\today \\
}

\vfill

\flushright{
\includegraphics[width=2.in]{../../../Bibliography/nistident_flright_vec} \\[.3in]
}

\titlesigs

\end{minipage}

\newpage

\frontmatter

\pagestyle{plain}
\pagenumbering{roman}

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{Contents}
\tableofcontents

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{List of Tables}
\listoftables

\chapter{List of Acronyms}

\begin{tabbing}
\hspace{1.5in} \= \\
FDS \> Fire Dynamics Simulator \\
HGL \> Hot Gas Layer \\
HRR \> Heat Release Rate \\
HRRPUA \> Heat Release Rate per Unit Area \\
NIST \> National Institute of Standards and Technology \\
\end{tabbing}

\mainmatter

\chapter{Introduction}
\label{chap:Introduction}

\chapter{Experimental Setup}
\label{chap:Experimental_Setup}

\section{Experimental Structure}
\label{sec:Experimental_Structure}

\subsection{Construction}
\label{sec:Construction}

\section{Fuel Load}

\chapter{Results}
\label{chap:Results}

\clearpage

\chapter{Conclusions}
\label{chap:Conclusions}

\chapter{Acknowledgements}
\label{chap:Acknowledgements}

\bibliography{../../../Bibliography/FDS_refs,../../../Bibliography/FDS_general,references}

\appendix

\chapter{All Tests}

\section{286 College}

\subsection{Test 1, Living room fire with sprinkler}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_1_TC_A3}
\caption{286 College, Test 1, Temperature TC\_A3}
\label{fig:286_College_Test_1_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_1_TC_A4}
\caption{286 College, Test 1, Temperature TC\_A4}
\label{fig:286_College_Test_1_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_1_TC_S_1}
\caption{286 College, Test 1, Temperature TC\_S\_1}
\label{fig:286_College_Test_1_TC_S_1}
\end{figure}

\clearpage

\subsection{Test 2, Garage fire, exterior attack}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_TC_A1}
\caption{286 College, Test 2, Temperature TC\_A1}
\label{fig:286_College_Test_2_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_TC_A2}
\caption{286 College, Test 2, Temperature TC\_A2}
\label{fig:286_College_Test_2_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_TC_A6}
\caption{286 College, Test 2, Temperature TC\_A6}
\label{fig:286_College_Test_2_TC_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_TC_A7}
\caption{286 College, Test 2, Temperature TC\_A7}
\label{fig:286_College_Test_2_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_TC_A8}
\caption{286 College, Test 2, Temperature TC\_A8}
\label{fig:286_College_Test_2_TC_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_HF}
\caption{286 College, Test 2, Heat Flux}
\label{fig:286_College_Test_2_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_P_S_1}
\caption{286 College, Test 2, Pressure}
\label{fig:286_College_Test_2_P_S_1}
\end{figure}

\clearpage

\subsection{Test 3, Living room fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_3_TC_A1}
\caption{286 College, Test 3, Temperature TC\_A1}
\label{fig:286_College_Test_3_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_3_TC_A4}
\caption{286 College, Test 3, Temperature TC\_A4}
\label{fig:286_College_Test_3_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_3_TC_A5}
\caption{286 College, Test 3, Temperature TC\_A5}
\label{fig:286_College_Test_3_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_3_HF}
\caption{286 College, Test 3, Heat Flux}
\label{fig:286_College_Test_3_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_3_P_S_1}
\caption{286 College, Test 3, Pressure}
\label{fig:286_College_Test_3_P_S_1}
\end{figure}

\clearpage

\section{489 N. Forest}

\subsection{Test 1, Living room fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_TC_A1}
\caption{489 N. Forest, Test 1, Temperature TC\_A1}
\label{fig:489_N_Forest_Test_1_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_TC_A2}
\caption{489 N. Forest, Test 1, Temperature TC\_A2}
\label{fig:489_N_Forest_Test_1_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_TC_A3}
\caption{489 N. Forest, Test 1, Temperature TC\_A3}
\label{fig:489_N_Forest_Test_1_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_TC_S_1}
\caption{489 N. Forest, Test 1, Temperature TC\_S\_1}
\label{fig:489_N_Forest_Test_1_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_HF}
\caption{489 N. Forest, Test 1, Heat Flux}
\label{fig:489_N_Forest_Test_1_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_P_S_1}
\caption{489 N. Forest, Test 1, Pressure}
\label{fig:489_N_Forest_Test_1_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_GAS}
\caption{489 N. Forest, Test 1, Gas Concentration}
\label{fig:489_N_Forest_Test_1_GAS}
\end{figure}

\clearpage

\subsection{Test 2, Second floor bedroom fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_TC_A4}
\caption{489 N. Forest, Test 2, Temperature TC\_A4}
\label{fig:489_N_Forest_Test_2_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_TC_A5}
\caption{489 N. Forest, Test 2, Temperature TC\_A5}
\label{fig:489_N_Forest_Test_2_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_TC_A6}
\caption{489 N. Forest, Test 2, Temperature TC\_A6}
\label{fig:489_N_Forest_Test_2_TC_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_TC_A7}
\caption{489 N. Forest, Test 2, Temperature TC\_A7}
\label{fig:489_N_Forest_Test_2_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_TC_S_1}
\caption{489 N. Forest, Test 2, Temperature TC\_S\_1}
\label{fig:489_N_Forest_Test_2_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_HF}
\caption{489 N. Forest, Test 2, Heat Flux}
\label{fig:489_N_Forest_Test_2_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_P_S_1}
\caption{489 N. Forest, Test 2, Pressure}
\label{fig:489_N_Forest_Test_2_P_S_1}
\end{figure}

\clearpage

\section{302 College}

\subsection{Test 1, Bedroom fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_TC_A1}
\caption{302 College, Test 1, Temperature TC\_A1}
\label{fig:302_College_Test_1_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_TC_A3}
\caption{302 College, Test 1, Temperature TC\_A3}
\label{fig:302_College_Test_1_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_TC_A4}
\caption{302 College, Test 1, Temperature TC\_A4}
\label{fig:302_College_Test_1_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_TC_A5}
\caption{302 College, Test 1, Temperature TC\_A5}
\label{fig:302_College_Test_1_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_TC_S_1}
\caption{302 College, Test 1, Temperature TC\_S\_1}
\label{fig:302_College_Test_1_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_HF}
\caption{302 College, Test 1, Heat Flux}
\label{fig:302_College_Test_1_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_P_S_1}
\caption{302 College, Test 1, Pressure}
\label{fig:302_College_Test_1_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_GAS}
\caption{302 College, Test 1, Gas Concentration}
\label{fig:302_College_Test_1_GAS}
\end{figure}

\clearpage

\subsection{Test 2, Living room fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_TC_A1}
\caption{302 College, Test 2, Temperature TC\_A1}
\label{fig:302_College_Test_2_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_TC_A3}
\caption{302 College, Test 2, Temperature TC\_A3}
\label{fig:302_College_Test_2_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_TC_A4}
\caption{302 College, Test 2, Temperature TC\_A4}
\label{fig:302_College_Test_2_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_TC_A5}
\caption{302 College, Test 2, Temperature TC\_A5}
\label{fig:302_College_Test_2_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_TC_S_1}
\caption{302 College, Test 2, Temperature TC\_S\_1}
\label{fig:302_College_Test_2_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_HF}
\caption{302 College, Test 2, Heat Flux}
\label{fig:302_College_Test_2_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_P_S_1}
\caption{302 College, Test 2, Pressure}
\label{fig:302_College_Test_2_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_GAS}
\caption{302 College, Test 2, Gas Concentration}
\label{fig:302_College_Test_2_GAS}
\end{figure}

\clearpage

\subsection{Test 3, Bathroom and attic fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_TC_A1}
\caption{302 College, Test 3, Temperature TC\_A1}
\label{fig:302_College_Test_3_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_TC_A2}
\caption{302 College, Test 3, Temperature TC\_A2}
\label{fig:302_College_Test_3_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_TC_A3}
\caption{302 College, Test 3, Temperature TC\_A3}
\label{fig:302_College_Test_3_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_TC_A4}
\caption{302 College, Test 3, Temperature TC\_A4}
\label{fig:302_College_Test_3_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_TC_A5}
\caption{302 College, Test 3, Temperature TC\_A5}
\label{fig:302_College_Test_3_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_TC_S_1}
\caption{302 College, Test 3, Temperature TC\_S\_1}
\label{fig:302_College_Test_3_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_HF}
\caption{302 College, Test 3, Heat Flux}
\label{fig:302_College_Test_3_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_P_S_1}
\caption{302 College, Test 3, Pressure}
\label{fig:302_College_Test_3_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_GAS}
\caption{302 College, Test 3, Gas Concentration}
\label{fig:302_College_Test_3_GAS}
\end{figure}

\clearpage

\section{305 Manning}

\subsection{Test 1, Garage fire, interior attack}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A1}
\caption{305 Manning, Test 1, Temperature TC\_A1}
\label{fig:305_Manning_Test_1_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A2}
\caption{305 Manning, Test 1, Temperature TC\_A2}
\label{fig:305_Manning_Test_1_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A3}
\caption{305 Manning, Test 1, Temperature TC\_A3}
\label{fig:305_Manning_Test_1_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A4}
\caption{305 Manning, Test 1, Temperature TC\_A4}
\label{fig:305_Manning_Test_1_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A5}
\caption{305 Manning, Test 1, Temperature TC\_A5}
\label{fig:305_Manning_Test_1_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A6}
\caption{305 Manning, Test 1, Temperature TC\_A6}
\label{fig:305_Manning_Test_1_TC_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A7}
\caption{305 Manning, Test 1, Temperature TC\_A7}
\label{fig:305_Manning_Test_1_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_HF}
\caption{305 Manning, Test 1, Heat Flux}
\label{fig:305_Manning_Test_1_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_P_S_1}
\caption{305 Manning, Test 1, Pressure}
\label{fig:305_Manning_Test_1_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_GAS}
\caption{305 Manning, Test 1, Gas Concentration}
\label{fig:305_Manning_Test_1_GAS}
\end{figure}

\clearpage

\subsection{Test 2, Bedroom fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A1}
\caption{305 Manning, Test 2, Temperature TC\_A1}
\label{fig:305_Manning_Test_2_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A2}
\caption{305 Manning, Test 2, Temperature TC\_A2}
\label{fig:305_Manning_Test_2_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A3}
\caption{305 Manning, Test 2, Temperature TC\_A3}
\label{fig:305_Manning_Test_2_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A4}
\caption{305 Manning, Test 2, Temperature TC\_A4}
\label{fig:305_Manning_Test_2_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A5}
\caption{305 Manning, Test 2, Temperature TC\_A5}
\label{fig:305_Manning_Test_2_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A6}
\caption{305 Manning, Test 2, Temperature TC\_A6}
\label{fig:305_Manning_Test_2_TC_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A7}
\caption{305 Manning, Test 2, Temperature TC\_A7}
\label{fig:305_Manning_Test_2_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_S_1}
\caption{305 Manning, Test 2, Temperature TC\_S\_1}
\label{fig:305_Manning_Test_2_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_HF}
\caption{305 Manning, Test 2, Heat Flux}
\label{fig:305_Manning_Test_2_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_P_S_1}
\caption{305 Manning, Test 2, Pressure}
\label{fig:305_Manning_Test_2_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_GAS}
\caption{305 Manning, Test 2, Gas Concentration}
\label{fig:305_Manning_Test_2_GAS}
\end{figure}

\clearpage

\subsection{Test 3, Complete burndown}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_3_HF}
\caption{305 Manning, Test 3, Heat Flux}
\label{fig:305_Manning_Test_3_HF}
\end{figure}

\clearpage

\section{289 Manning}

\subsection{Test 1, Basement fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_TC_A1}
\caption{286 College, Test 1, Temperature TC\_A1}
\label{fig:289_Manning_Test_1_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_TC_A4}
\caption{286 College, Test 1, Temperature TC\_A4}
\label{fig:289_Manning_Test_1_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_TC_A5}
\caption{286 College, Test 1, Temperature TC\_A5}
\label{fig:289_Manning_Test_1_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_TC_A6}
\caption{286 College, Test 1, Temperature TC\_A6}
\label{fig:289_Manning_Test_1_TC_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_TC_A7}
\caption{286 College, Test 1, Temperature TC\_A7}
\label{fig:289_Manning_Test_1_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_TC_S_1}
\caption{286 College, Test 1, Temperature TC\_S\_1}
\label{fig:289_Manning_Test_1_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_HF}
\caption{286 College, Test 1, Heat Flux}
\label{fig:289_Manning_Test_1_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_P_S_1}
\caption{286 College, Test 1, Pressure}
\label{fig:289_Manning_Test_1_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_GAS}
\caption{286 College, Test 1, Gas Concentration}
\label{fig:289_Manning_Test_1_GAS}
\end{figure}

\clearpage

\subsection{Test 2, Deck and attic fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_2_TC_A2}
\caption{286 College, Test 2, Temperature TC\_A2}
\label{fig:289_Manning_Test_2_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_2_TC_A3}
\caption{286 College, Test 2, Temperature TC\_A3}
\label{fig:289_Manning_Test_2_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_2_HF}
\caption{286 College, Test 2, Heat Flux}
\label{fig:289_Manning_Test_2_HF}
\end{figure}

\end{document}