### 2014-08-27 Projects/Spartanburg_2014/Report/Spartanburg_2014.tex
Spartanburg 2014: Add 286 College Test Descriptions to report appendix

\documentclass[12pt,oneside]{book}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                                                 %
% The mathematical style of these documents follows                                               %
%                                                                                                 %
% A. Thompson and B.N. Taylor. The NIST Guide for the Use of the International System of Units.   %
%    NIST Special Publication 881, 2008.                                                          %
%                                                                                                 %
% http://www.nist.gov/pml/pubs/sp811/index.cfm                                                    %
%                                                                                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{../../../Bibliography/commoncommands}

% Rename chapter headings
\renewcommand{\chaptername}{Section}
\renewcommand{\bibname}{References}

% Math shortcuts
\renewcommand{\sb}[1]{_\mathrm{#1}}
\renewcommand{\C}{\mbox{C}}
\renewcommand{\H}{\mbox{H}}
\renewcommand{\O}{\mbox{O}}
\newcommand{\N}{\mbox{N}}

% Center all figures
\makeatletter
\g@addto@macro\@floatboxreset\centering
\makeatother

% Extra packages
\usepackage{xfrac}

\begin{document}

\bibliographystyle{unsrt}
\pagestyle{empty}

\begin{minipage}[t][9in][s]{6.25in}

\begin{flushright}
\fontsize{20}{24}\selectfont
\bf{NIST Technical Note XXXX}
\end{flushright}

\headerB{
Spartanburg Fire Tests \\
2013--2014 \\
}

\normalsize

\headerC{
{
\flushright{
Daniel Madrzykowski \\
Kristopher J. Overholt \\
Craig G. Weinschenk \\
Keith M. Stakes \\

\vspace*{2\baselineskip}

\begingroup
\hypersetup{urlcolor=black}
\href{http://dx.doi.org/10.6028/NIST.TN.XXXX}{http://dx.doi.org/10.6028/NIST.TN.XXXX}
\endgroup
}

\vfill

\flushright{

\includegraphics[width=2.in]{../../../Bibliography/nistident_flright_vec} \\[.3in]
}
}
}

\end{minipage}

\newpage
\hspace{5in}
\newpage

\frontmatter

\pagenumbering{roman}

\begin{minipage}[t][9in][s]{6.25in}

\begin{flushright}
\fontsize{20}{24}\selectfont
\bf{NIST Technical Note XXXX}
\end{flushright}

\headerB{
Spartanburg Fire Tests \\
2013--2014 \\
}

\headerC{
\flushright{
Daniel Madrzykowski \\
Kristopher J. Overholt \\
Craig G. Weinschenk \\
Keith M. Stakes \\
{\em Fire Research Division \\
Engineering Laboratory} \\

\vspace*{2\baselineskip}

\begingroup
\hypersetup{urlcolor=black}
\href{http://dx.doi.org/10.6028/NIST.TN.XXXX}{http://dx.doi.org/10.6028/NIST.TN.XXXX} \\
\endgroup

\vspace*{2\baselineskip}

August 2014}}

\vfill

\flushright{\includegraphics[width=1in]{../../../Bibliography/doc} }

\titlesigs

\end{minipage}

\newpage

\begin{minipage}[t][9in][s]{6.25in}

\flushright{Certain commercial entities, equipment, or materials may be identified in this \\
document in order to describe an experimental procedure or concept adequately. \\
Such identification is not intended to imply recommendation or endorsement by the \\
National Institute of Standards and Technology, nor is it intended to imply that the \\
entities, materials, or equipment are necessarily the best available for the purpose. \\
}

\vspace{3in}

\large
\flushright{\bf National Institute of Standards and Technology Technical Note XXXX \\
Natl.~Inst.~Stand.~Technol.~Tech.~Note~XX, \pageref{LastPage} pages (August 2014) \\
% http://dx.doi.org/10.6028/NIST.TN.XXXX \\
CODEN: NTNOEF }

\vfill

\hspace{1in}

\end{minipage}

\newpage

\frontmatter

\pagestyle{plain}
\pagenumbering{roman}

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{Contents}
\tableofcontents

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{List of Tables}
\listoftables

\chapter{List of Acronyms}

\begin{tabbing}
\hspace{1.5in} \= \\
FDS \> Fire Dynamics Simulator \\
HGL \> Hot Gas Layer \\
HRR \> Heat Release Rate \\
HRRPUA \> Heat Release Rate per Unit Area \\
NIST \> National Institute of Standards and Technology \\
\end{tabbing}

\mainmatter

\chapter{Introduction}
\label{chap:Introduction}

\chapter{Experimental Setup}
\label{chap:Experimental_Setup}

\section{Experimental Structure}
\label{sec:Experimental_Structure}

\section{Construction}
\label{sec:Construction}

\section{Instrumentation}
\label{sec:Instrumentation}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Instrumentation/286_College}
\caption{Instrumentation, 286 College}
\label{fig:Instrumentation_286_College}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Instrumentation/489_N_Forest_1st_Floor}
\caption{Instrumentation, 489 N. Forest, first floor}
\label{fig:Instrumentation_489_N_Forest_1st_Floor}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Instrumentation/489_N_Forest_2nd_Floor}
\caption{Instrumentation, 489 N. Forest, second floor}
\label{fig:Instrumentation_489_N_Forest_2nd_Floor}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Instrumentation/302_College}
\caption{Instrumentation, 302 College}
\label{fig:Instrumentation_302_College}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Instrumentation/305_Manning}
\caption{Instrumentation, 305 Manning}
\label{fig:Instrumentation_305_Manning}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Instrumentation/289_Manning_1st_Floor}
\caption{Instrumentation, 289 Manning, first floor}
\label{fig:Instrumentation_289_Manning_1st_Floor}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=6in]{../Figures/Instrumentation/289_Manning_Basement}
\caption{Instrumentation, 289 Manning, basement}
\label{fig:Instrumentation_289_Manning_basement}
\end{figure}

\clearpage

\section{Fuel Load}
\label{sec:Fuel_Load}

\begin{sidewaystable}[!ht]
\centering
\caption{Fuel masses}
\begin{tabular}{llcc}
\hline\noalign{\smallskip}
Item                         &  Material Description             &  Dimensions (m)            &  Mass (kg)  \\
\noalign{\smallskip}\hline\noalign{\smallskip}
% 3-seat Legacy Sofa         &  491 Brawley                      &  1.96 m x 0.86 m x 0.89 m  &  56.2 kg    \\
%   Cushion (left)           &  Cotton w/ inner springs          &                            &  4.06 kg    \\
%   Cushions (center)        &  Cotton w/ inner springs          &                            &  4.56 kg    \\
%   Cushions (rear)          &  Cotton w/ inner springs          &                            &  4.16 kg    \\
% 2-seat New Sofa            &  491 Brawley                      &  1.59 m x 0.91 m x 0.86 m  &  33.3 kg    \\
%   Cushion (seat)           &                                   &  0.56 m x 0.61 m x 0.14 m  &  1.6 kg     \\
%   Cushion (rear)           &  0.71 m at top, 0.58 m at bottom  &  Varies x 0.53 m x 0.20 m  &  2.0 kg     \\
3-seat Brown Sofa            &  302 College (front room)         &  2.24 m x 1.02 m x 0.91 m  &  61.0 kg    \\
2-seat Purple Stripe Sofa    &  302 College (front room)         &  1.57 m x 0.97 m x 0.86 m  &  82.0 kg    \\
3-seat Black Pine Leaf Sofa  &  302 College (front room)         &  2.08 m x 0.86 m x 0.86 m  &  51.0 kg    \\
Blue/Brown Chair             &  302 College (front room)         &  0.97 m x 0.81 m x 0.97 m  &  46.4 kg    \\
2-seat Floral Sofa           &  302 College (rear room)          &  1.68 m x 0.91 m x 0.91 m  &  34.5 kg    \\
2-seat Blue Plaid Sofa       &  302 College (rear room)          &  1.63 m x 0.97 m x 0.91 m  &  39.8 kg    \\
3-seat Yellow Sofa           &  302 College (rear room)          &  2.13 m x 0.91 m x 0.71 m  &  48.3 kg    \\
3-seat Floral Sofa           &  302 College (rear room)          &  2.29 m x 0.91 m x 0.91 m  &  48.0 kg    \\
3-seat New Sofa              &  289 College (front room)         &  2.13 m x 0.91 m x 0.91 m  &  43.0 kg    \\
Blue Chair                   &  289 College (middle room)        &  0.91 m x 0.91 m x 0.91 m  &  25.3 kg    \\
3-seat Floral Sofa           &  289 College (sprinkler room)     &  2.24 m x 1.02 m x 1.02 m  &  81.2 kg    \\
Floral Chair                 &  289 College (sprinkler room)     &  1.02 m x 1.02 m x 1.02 m  &  36.3 kg    \\
\noalign{\smallskip}\hline
\end{tabular}
\label{tab:Fuel_Masses}
\end{sidewaystable}

\chapter{Results}
\label{chap:Results}

\clearpage

\chapter{Conclusions}
\label{chap:Conclusions}

\chapter{Acknowledgments}
\label{chap:Acknowledgments}

\bibliography{../../../Bibliography/FDS_refs,../../../Bibliography/FDS_general,references}

\appendix

\chapter{All Tests}

\section{286 College}

The first set of experiments were conducted in a single story residential structure with attached garage addressed as 286 College Street.  The structure consisted of type 5 construction. The attached garage was added on to the B side of the structure for the purpose of these tests.  The non-burn rooms had a limited-combustible finish comprised of either paper-covered gypsum board or a finished plaster board.  The burn rooms had a combustible finish comprised of OSB sheets for wall linings.  Three tests were conducted in this structure: 1) A living room fire with a residential sprinkler present, 2) An attached garage fire with exterior attack, and 3) A living room fire with door control.

\subsection{Test 1, Living room fire with sprinkler}

Test 1 consisted of a fire in the living room with residential sprinkler activation.  Ignition occurred on a couch in the living room to the rear of the structure near the C/D corner.  The fire was remotely ignited from the exterior of the structure.  There was a residential sprinkler in the fire room.  Other fire room furnishings included a sofa chair, two wooden end tables, a floor lamp, cloth window curtains, carpeted floor, and wooden paneling for wall linings.  The fire grew uninhibited until the sprinkler activated and suppressed the bulk of the fire.  A fire fighting crew made entry through side C for mop-up.

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_1_TC_A3}
\caption{286 College, Test 1, Temperature TC\_A3}
\label{fig:286_College_Test_1_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_1_TC_A4}
\caption{286 College, Test 1, Temperature TC\_A4}
\label{fig:286_College_Test_1_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_1_TC_S_1}
\caption{286 College, Test 1, Temperature TC\_S\_1}
\label{fig:286_College_Test_1_TC_S_1}
\end{figure}

\clearpage

\subsection{Test 2, Garage fire, exterior attack}

Test 2 consisted of a fire in the attached garage with an exterior fire attack.  Two separate ignition points were used, one of which was inside a car located in the garage and another was located on a couch to the rear of the garage.  The couch simulated other fuel loads commonly found within a garage.  Several wooden pallets were also placed in the garage for additional fuel.  The fire was ignited remotely from the exterior of the garage.  The door from the attached garage to the interior of the structure (kitchen area) remained closed for the duration of this test.  The fire grew uninhibited in the garage until the space reached flashover conditions and became fully-involved.  The fire was knocked down and suppressed from the exterior of the structure by a fire fighting crew with a single hand-line in a straight stream pattern.  The front door on Side A of the structure was then opened to simulate a crew making entry.

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_TC_A1}
\caption{286 College, Test 2, Temperature TC\_A1}
\label{fig:286_College_Test_2_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_TC_A2}
\caption{286 College, Test 2, Temperature TC\_A2}
\label{fig:286_College_Test_2_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_TC_A6}
\caption{286 College, Test 2, Temperature TC\_A6}
\label{fig:286_College_Test_2_TC_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_TC_A7}
\caption{286 College, Test 2, Temperature TC\_A7}
\label{fig:286_College_Test_2_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_TC_A8}
\caption{286 College, Test 2, Temperature TC\_A8}
\label{fig:286_College_Test_2_TC_A8}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_HF}
\caption{286 College, Test 2, Heat Flux}
\label{fig:286_College_Test_2_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_2_P_S_1}
\caption{286 College, Test 2, Pressure}
\label{fig:286_College_Test_2_P_S_1}
\end{figure}

\clearpage

\subsection{Test 3, Living room fire}

Test 3 consisted of a fire in the living room located in the front of the structure near the A/B corner.  The ignition point for this test was the corner of a modern-style couch and was performed remotely from the exterior of the structure.  Other fuels present in the room were another modern-style sofa and a modern-style love seat.  The fire room was lined with OSB paneling for additional fuel.  The front side A door was in the open position for the start of the test.  The target room, located adjacent to the fire room near the A/D corner, was separated by a closed hollow core wooden door.  Once ignited, the fire grew uninhibited until the room approached flashover conditions.  Prior to this, the front door to the room was closed.  The fire was allowed to stabilize and then the front door was opened.  The fire fighting crew made an attack from the exterior via a straight stream pattern from a single hand-line.  The pattern was then switched to a narrow fog for further suppression.  The fire fighting crew then made entry into the fire room for mop-up procedures.

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_3_TC_A1}
\caption{286 College, Test 3, Temperature TC\_A1}
\label{fig:286_College_Test_3_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_3_TC_A4}
\caption{286 College, Test 3, Temperature TC\_A4}
\label{fig:286_College_Test_3_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_3_TC_A5}
\caption{286 College, Test 3, Temperature TC\_A5}
\label{fig:286_College_Test_3_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_3_HF}
\caption{286 College, Test 3, Heat Flux}
\label{fig:286_College_Test_3_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/286_College_Test_3_P_S_1}
\caption{286 College, Test 3, Pressure}
\label{fig:286_College_Test_3_P_S_1}
\end{figure}

\clearpage

\section{489 N. Forest}

\subsection{Test 1, Living room fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_TC_A1}
\caption{489 N. Forest, Test 1, Temperature TC\_A1}
\label{fig:489_N_Forest_Test_1_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_TC_A2}
\caption{489 N. Forest, Test 1, Temperature TC\_A2}
\label{fig:489_N_Forest_Test_1_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_TC_A3}
\caption{489 N. Forest, Test 1, Temperature TC\_A3}
\label{fig:489_N_Forest_Test_1_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_TC_S_1}
\caption{489 N. Forest, Test 1, Temperature TC\_S\_1}
\label{fig:489_N_Forest_Test_1_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_HF}
\caption{489 N. Forest, Test 1, Heat Flux}
\label{fig:489_N_Forest_Test_1_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_P_S_1}
\caption{489 N. Forest, Test 1, Pressure}
\label{fig:489_N_Forest_Test_1_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_1_GAS}
\caption{489 N. Forest, Test 1, Gas Concentration}
\label{fig:489_N_Forest_Test_1_GAS}
\end{figure}

\clearpage

\subsection{Test 2, Second floor bedroom fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_TC_A4}
\caption{489 N. Forest, Test 2, Temperature TC\_A4}
\label{fig:489_N_Forest_Test_2_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_TC_A5}
\caption{489 N. Forest, Test 2, Temperature TC\_A5}
\label{fig:489_N_Forest_Test_2_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_TC_A6}
\caption{489 N. Forest, Test 2, Temperature TC\_A6}
\label{fig:489_N_Forest_Test_2_TC_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_TC_A7}
\caption{489 N. Forest, Test 2, Temperature TC\_A7}
\label{fig:489_N_Forest_Test_2_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_TC_S_1}
\caption{489 N. Forest, Test 2, Temperature TC\_S\_1}
\label{fig:489_N_Forest_Test_2_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_HF}
\caption{489 N. Forest, Test 2, Heat Flux}
\label{fig:489_N_Forest_Test_2_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/489_N_Forest_Test_2_P_S_1}
\caption{489 N. Forest, Test 2, Pressure}
\label{fig:489_N_Forest_Test_2_P_S_1}
\end{figure}

\clearpage

\section{302 College}

\subsection{Test 1, Bedroom fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_TC_A1}
\caption{302 College, Test 1, Temperature TC\_A1}
\label{fig:302_College_Test_1_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_TC_A3}
\caption{302 College, Test 1, Temperature TC\_A3}
\label{fig:302_College_Test_1_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_TC_A4}
\caption{302 College, Test 1, Temperature TC\_A4}
\label{fig:302_College_Test_1_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_TC_A5}
\caption{302 College, Test 1, Temperature TC\_A5}
\label{fig:302_College_Test_1_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_TC_S_1}
\caption{302 College, Test 1, Temperature TC\_S\_1}
\label{fig:302_College_Test_1_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_HF}
\caption{302 College, Test 1, Heat Flux}
\label{fig:302_College_Test_1_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_P_S_1}
\caption{302 College, Test 1, Pressure}
\label{fig:302_College_Test_1_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_1_GAS}
\caption{302 College, Test 1, Gas Concentration}
\label{fig:302_College_Test_1_GAS}
\end{figure}

\clearpage

\subsection{Test 2, Living room fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_TC_A1}
\caption{302 College, Test 2, Temperature TC\_A1}
\label{fig:302_College_Test_2_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_TC_A3}
\caption{302 College, Test 2, Temperature TC\_A3}
\label{fig:302_College_Test_2_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_TC_A4}
\caption{302 College, Test 2, Temperature TC\_A4}
\label{fig:302_College_Test_2_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_TC_A5}
\caption{302 College, Test 2, Temperature TC\_A5}
\label{fig:302_College_Test_2_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_TC_S_1}
\caption{302 College, Test 2, Temperature TC\_S\_1}
\label{fig:302_College_Test_2_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_HF}
\caption{302 College, Test 2, Heat Flux}
\label{fig:302_College_Test_2_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_P_S_1}
\caption{302 College, Test 2, Pressure}
\label{fig:302_College_Test_2_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_2_GAS}
\caption{302 College, Test 2, Gas Concentration}
\label{fig:302_College_Test_2_GAS}
\end{figure}

\clearpage

\subsection{Test 3, Bathroom and attic fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_TC_A1}
\caption{302 College, Test 3, Temperature TC\_A1}
\label{fig:302_College_Test_3_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_TC_A2}
\caption{302 College, Test 3, Temperature TC\_A2}
\label{fig:302_College_Test_3_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_TC_A3}
\caption{302 College, Test 3, Temperature TC\_A3}
\label{fig:302_College_Test_3_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_TC_A4}
\caption{302 College, Test 3, Temperature TC\_A4}
\label{fig:302_College_Test_3_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_TC_A5}
\caption{302 College, Test 3, Temperature TC\_A5}
\label{fig:302_College_Test_3_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_TC_S_1}
\caption{302 College, Test 3, Temperature TC\_S\_1}
\label{fig:302_College_Test_3_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_HF}
\caption{302 College, Test 3, Heat Flux}
\label{fig:302_College_Test_3_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_P_S_1}
\caption{302 College, Test 3, Pressure}
\label{fig:302_College_Test_3_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/302_College_Test_3_GAS}
\caption{302 College, Test 3, Gas Concentration}
\label{fig:302_College_Test_3_GAS}
\end{figure}

\clearpage

\section{305 Manning}

\subsection{Test 1, Garage fire, interior attack}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A1}
\caption{305 Manning, Test 1, Temperature TC\_A1}
\label{fig:305_Manning_Test_1_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A2}
\caption{305 Manning, Test 1, Temperature TC\_A2}
\label{fig:305_Manning_Test_1_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A3}
\caption{305 Manning, Test 1, Temperature TC\_A3}
\label{fig:305_Manning_Test_1_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A4}
\caption{305 Manning, Test 1, Temperature TC\_A4}
\label{fig:305_Manning_Test_1_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A5}
\caption{305 Manning, Test 1, Temperature TC\_A5}
\label{fig:305_Manning_Test_1_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A6}
\caption{305 Manning, Test 1, Temperature TC\_A6}
\label{fig:305_Manning_Test_1_TC_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_TC_A7}
\caption{305 Manning, Test 1, Temperature TC\_A7}
\label{fig:305_Manning_Test_1_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_HF}
\caption{305 Manning, Test 1, Heat Flux}
\label{fig:305_Manning_Test_1_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_P_S_1}
\caption{305 Manning, Test 1, Pressure}
\label{fig:305_Manning_Test_1_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_1_GAS}
\caption{305 Manning, Test 1, Gas Concentration}
\label{fig:305_Manning_Test_1_GAS}
\end{figure}

\clearpage

\subsection{Test 2, Bedroom fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A1}
\caption{305 Manning, Test 2, Temperature TC\_A1}
\label{fig:305_Manning_Test_2_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A2}
\caption{305 Manning, Test 2, Temperature TC\_A2}
\label{fig:305_Manning_Test_2_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A3}
\caption{305 Manning, Test 2, Temperature TC\_A3}
\label{fig:305_Manning_Test_2_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A4}
\caption{305 Manning, Test 2, Temperature TC\_A4}
\label{fig:305_Manning_Test_2_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A5}
\caption{305 Manning, Test 2, Temperature TC\_A5}
\label{fig:305_Manning_Test_2_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A6}
\caption{305 Manning, Test 2, Temperature TC\_A6}
\label{fig:305_Manning_Test_2_TC_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_A7}
\caption{305 Manning, Test 2, Temperature TC\_A7}
\label{fig:305_Manning_Test_2_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_TC_S_1}
\caption{305 Manning, Test 2, Temperature TC\_S\_1}
\label{fig:305_Manning_Test_2_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_HF}
\caption{305 Manning, Test 2, Heat Flux}
\label{fig:305_Manning_Test_2_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_P_S_1}
\caption{305 Manning, Test 2, Pressure}
\label{fig:305_Manning_Test_2_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_2_GAS}
\caption{305 Manning, Test 2, Gas Concentration}
\label{fig:305_Manning_Test_2_GAS}
\end{figure}

% \clearpage

% Heat flux measurements are large negative values

% \subsection{Test 3, Complete burndown}

% \begin{figure}[!ht]
% \includegraphics[width=5in]{../Figures/Script_Figures/305_Manning_Test_3_HF}
% \caption{305 Manning, Test 3, Heat Flux}
% \label{fig:305_Manning_Test_3_HF}
% \end{figure}

\clearpage

\section{289 Manning}

\subsection{Test 1, Basement fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_TC_A1}
\caption{286 College, Test 1, Temperature TC\_A1}
\label{fig:289_Manning_Test_1_TC_A1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_TC_A4}
\caption{286 College, Test 1, Temperature TC\_A4}
\label{fig:289_Manning_Test_1_TC_A4}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_TC_A5}
\caption{286 College, Test 1, Temperature TC\_A5}
\label{fig:289_Manning_Test_1_TC_A5}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_TC_A6}
\caption{286 College, Test 1, Temperature TC\_A6}
\label{fig:289_Manning_Test_1_TC_A6}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_TC_A7}
\caption{286 College, Test 1, Temperature TC\_A7}
\label{fig:289_Manning_Test_1_TC_A7}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_TC_S_1}
\caption{286 College, Test 1, Temperature TC\_S\_1}
\label{fig:289_Manning_Test_1_TC_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_HF}
\caption{286 College, Test 1, Heat Flux}
\label{fig:289_Manning_Test_1_HF}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_P_S_1}
\caption{286 College, Test 1, Pressure}
\label{fig:289_Manning_Test_1_P_S_1}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_1_GAS}
\caption{286 College, Test 1, Gas Concentration}
\label{fig:289_Manning_Test_1_GAS}
\end{figure}

\clearpage

\subsection{Test 2, Deck and attic fire}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_2_TC_A2}
\caption{286 College, Test 2, Temperature TC\_A2}
\label{fig:289_Manning_Test_2_TC_A2}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_2_TC_A3}
\caption{286 College, Test 2, Temperature TC\_A3}
\label{fig:289_Manning_Test_2_TC_A3}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=5in]{../Figures/Script_Figures/289_Manning_Test_2_HF}
\caption{286 College, Test 2, Heat Flux}
\label{fig:289_Manning_Test_2_HF}
\end{figure}

\end{document}