### 2015-01-22 Projects/DelCo_2014/Report/Hose_Tests/DelCo_Hose_Tests_2014_Outline.tex
DelCo_2014: Created outline for hose test report and edited hose test report

\documentclass{article}
\usepackage{outlines}
\usepackage{enumitem}
\setenumerate[1]{label=\Roman*.}
\setenumerate[2]{label=\Alph*.}
\setenumerate[3]{label=\roman*.}
\setenumerate[4]{label=\alph*.}
\begin{document}

DelCo 2014 Hose Stream Report Outline\\

\begin{outline}[enumerate]
\1 Introduction 
	\2 Background
		\3 Overview of different techniques used throughout the fire service
		\3 Brief review of hose stream research already in existence
	\2 Description of stream and application patterns considered in report

\1 Experimental Setup
	\2 Structures
		\3 General construction (picture of both structures)
		\3 East structure dimensioned floor plan
		\3 West structure dimensioned floor plan 
	\2 Instrumentation
		\3 Description/picture(s) of instrumentation used
		\3 East structure floor plan drawing with BDP/TC locations
		\3 West structure floor plan drawing with BDP/TC locations
	\2 Uncertainty
		\3 Explanation of Type A and B uncertainties
		\3 Length measurements
		\3 BDP velocity measurements
		\3 Water flow rate and hoseline pressure measurements

\1 Experimental Procedure
	\2 Cold flow tests
		\3 Differences in each procedure (e.g., different flow path configurations)
	\2 Water flow tests
		\3 Monitor aimed at various locations
		\3 Handline with various patterns
		\3 Differences in each procedure (e.g., different flow path configurations)

\1 Results
	\2 Cold flow tests
	\2 Water flow tests
		\3 Monitor at different locations
		\3 Hose stream patterns
		\3 Application patterns
			\4 No difference between CW and CCW

\1 Conclusion

\1 Future Work

\1 Acknowledgements

\1 References

\end{outline}

\end{document}