### 2015-07-10 Projects/DelCo_2015/Report/Test_Matrix.tex
DelCo_2015: add tex file with table of tests

\documentclass{article}

\usepackage{nopageno}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{pdfsync}
\usepackage{microtype}
\usepackage{siunitx}
\usepackage{rotating}
\usepackage{tabularx,ragged2e,booktabs,caption}
\usepackage{caption}
\frenchspacing

\begin{document}

\begin{sidewaystable}[htbp]
\centering
\captionof{table}{DelCo ESTC 2015 Tests}\label{tab:tests}
\begin{tabular}{lllll}
\toprule[1.5pt]
Test \#         &  Vent Configuration         	& Attack           & Hose Stream                         & Gear      \\
\midrule
30              &  Cherry Road                	& Uncoordinated    &  Straight, 120~gpm                  & 3 samples \\
31              &  Cherry Road; single basement & Uncoordinated    &  Straight, 120~gpm                  & 3 samples \\
32              &  Cherry Road + upper C doors  & Uncoordinated    &  Straight, 120~gpm                  & 2 samples \\
33              &  PPV Test                     & & &           \\  
34              &  Cherry Road                  & Coordinated      &  Straight, 120~gpm                  & 3 samples \\   
35              &  Cherry Road + PPV            & Uncoordinated    &  Straight, 180~gpm + interior fog   & 3 samples \\
36              &  Cherry Road                  & Coordinated      &  Straight, 180~gpm + interior fog   & 3 samples + manikin \\
37              &  Cherry Road                  & Uncoordinated    &  Straight, 180~gpm + interior fog   & 3 samples + manikin \\
38              &  Open A door + C doors        &                  &  Straight, CAFS, 120~gpm            & 3 samples + facepieces \\
39				&  Open A door + C doors        &                  &  Straight, 130~gpm                  & \\
40				&  Open A door + C doors        &                  &  Straight, CAFS, 130~gpm            & \\
41				&  Open A door + C doors        &                  &  CAFS, varied                       & \\
42				&  Cherry Road + upper C doors  & Allow decay      &  Straight, 150~gpm                  & \\
43				&  Open A door + C doors        &                  &  Straight/narrow fog, 120/180~gpm   & 1 sample \\
44				&  Open A door + C doors        &                  &  Interior wide fog, 130~gpm         & 1 sample, facepiece \\
45              &  Open A door + C doors        & Delay water 44   &  Interior wide fog, 130~gpm         & 1 sample, facepiece, helmet \\
45              &  Cherry Road; single basement & Uncoordinated    &  Straight, 150~gpm					 & \\
\bottomrule[1.25pt]
\end{tabular}
\end{sidewaystable}

\end{document}